### 2022-10-05 paper.qmd
one more

---
title: "Digitization of the Australian Parliamentary Debates (1901-2022)"
format: pdf
editor: visual
author:
  - name: Lindsay Katz
  - name: Rohan Alexander
---

# Introduction
- Mention the available formats of Hansard

# Existing Datasets
- What already exists, and how has it been used
- How are these existing data sets lacking?
- How do we account for this in our data set?
    - WhoGov article has good summary table offering comparison to other existing data sets

# Description of the Dataset
- I figure this would be a good spot to highlight everything within the dataset, as well as descriptive statistics

# Creation of the Dataset

## Interjections

## Question time

# Validation
- Might also mention strengths and weaknesses of our dataset

# Applications

# Conclusion

# References