### 2023-03-30 paper/paper-nature_format.tex

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  letterpaper,
  DIV=11,
  numbers=noendperiod]{scrartcl}

\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{5}
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\KOMAoption{captions}{tableheading}
\makeatletter
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[many]{tcolorbox}}
\makeatother
\makeatletter
\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
\makeatother
\makeatletter
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Digitization of the Australian Parliamentary Debates, 1998-2022},
  pdfauthor={Lindsay Katz; Rohan Alexander},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\title{Digitization of the Australian Parliamentary Debates, 1998-2022}
\author{Lindsay Katz\footnote{University of Toronto,
  lindsay.katz@mail.utoronto.ca} \and Rohan Alexander\footnote{University
  of Toronto, rohan.alexander@utoronto.ca}}
\date{March 30, 2023}

\begin{document}
\maketitle
\begin{abstract}
Public knowledge of what is said in parliament is a tenet of democracy,
and a critical resource for political science research. In Australia,
following the British tradition, the written record of what is said in
parliament is known as Hansard. While the Australian Hansard has always
been publicly available, it has been difficult to use for the purpose of
large-scale macro and micro-level text analysis because it has only been
available as PDFs or XMLs. Following the lead of the Linked
Parliamentary Data project which achieved this for Canada, we provide a
new, comprehensive, high-quality, rectangular database that captures all
proceedings of the Australian parliamentary debates from 1998 to the
present using Hansard. The database is publicly available and can be
linked to other datasets such as election results. The creation and
accessibility of this database enables the exploration of questions that
have not been possible to explore, serving as a valuable resource for
both researchers and policymakers.
\end{abstract}
\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[interior hidden, borderline west={3pt}{0pt}{shadecolor}, breakable, frame hidden, sharp corners, boxrule=0pt, enhanced]}{\end{tcolorbox}}\fi

\hypertarget{sec-intro}{%
\section{Background \& Summary}\label{sec-intro}}

The official written record of parliamentary debates, formally known as
Hansard, plays a fundamental role in capturing the history of political
proceedings and facilitating the exploration of valuable research
questions. Originating in the British parliament, the production of
Hansard became tradition in many Commonwealth countries such as Canada,
the United Kingdom, and Australia (Vice and Farrell 2017). Given the
content and magnitude of these records, they have significance,
particularly in the context of political science research. In the case
of Canada, the Hansard has been digitized for 1901 to 2019 (Beelen et
al. 2017). Having a digitized version of Hansard enables researchers to
perform advanced analyses on these records using text analysis tools and
statistical modelling. Following the lead of that project, in this paper
we introduce a novel database for the Australian Hansard. This is
composed of individual datasets for each sitting day in the House of
Representatives from 1998 to 2022, containing details on everything said
in parliament in a form that can be readily used by researchers. With
the development of tools for large-scale text analysis, this database
will serve as a valuable resource for efficiently studying and exploring
political behaviour in Australia over time.

The Australian House of Representatives performs a number of crucial
governmental functions, such as creating new laws and overseeing
government expenditure (House of Representatives 2018, ch.~1). The House
operates under a parallel chamber setup, meaning there are two debate
venues where proceedings take place: the Chamber, and the Federation
Chamber. Sittings of the House follow a predefined order of business,
regulated by procedural rules called standing orders (House of
Representatives 2018, ch.~8). A typical sitting day in the Chamber has a
number of scheduled proceedings including debates on government
business, 90 second member statements, and Question Time (House of
Representatives 2018, ch.~8). The Federation Chamber was created in 1994
as a subordinate debate venue of the Chamber. This allows for better
time management of House business as its proceedings occur
simultaneously with those of the Chamber (House of Representatives 2018,
ch.~21). Sittings in the Federation Chamber are quite different to those
of the Chamber in terms of their order of business and scope of
discussion. Business matters discussed in the Federation Chamber are
limited largely to intermediate stages of bill development, and the
business of private Members (House of Representatives 2018, ch.~21). It
is the recording and compilation of all of these proceedings on which
Hansard is based, and it is essentially, but not entirely, verbatim.

A week or so after each sitting day, a transcript is available for
download from the official Parliament of Australia website in both PDF
and extensible markup language (XML) form. The PDF is the official
release. For early decades there is a need to digitize PDFs using
Optical Character Recognition (OCR) technology and there are many flaws
(Sherratt 2016). The PDF imposes formatting designed for humans to read
with ease, whereas XML is designed for consistency and machine
legibility. The nature of XML enables us to more easily use code to
manipulate these records at scale, motivating our choice to develop our
database solely using the XML formatted files. In cases where we were
unsure on how to proceed with processing the XML, we defer first to the
PDF, and then to the video recording of the proceeding, if available.

At present, the Hansard format that is available on the Parliament of
Australia website is not accessible for large scale analysis. To this
point, various researchers have had to create their own databases of
usable, complete data based on content from the Australian Parliament
website. For instance, Sherratt (2016) created an online, easy to read
database of Hansard from 1901 to 1980 using the XML files. These data
can be navigated by year, parliament, people, and bills (Sherratt 2016).
To make the Australian Parliamentary Handbook more accessible, Leslie
(2021) has released an \texttt{R} package which includes data on all
Members of the Australian House of Representatives from 1945 to 2019.
Further, Alexander and Hodgetts (2021) created the
\texttt{AustralianPoliticians} \texttt{R} package, which contains
several datasets related to the political and biographical information
of Australian federal politicians who were active between 1901 and 2021.

Many papers exist which use components of Australian Hansard to explore
various research topics. For example, Salisbury (2011) used online
Hansard records to investigate the occurrences of unparliamentary
comments by members of the House, where the Speaker tells that member to
withdraw their remark. Rasiah (2010) worked with Question Time data from
Hansard transcripts during February and March of 2003, to investigate
resistance of politicians in answering questions about Iraq. In their
work, Fraussen, Graham, and Halpin (2018) use Hansard to quantify
political prominence by investigating strategic mentions of interest
groups by elected officials. Finally, the work of Alexander and
Alexander (2021) offers a dataset of the Australian Hansard, along with
an analysis of the effect of elections and changes in Prime Ministers
upon topics mentioned in parliament. Alexander and Alexander (2021)
created this database with the static PDF versions of Hansard, using OCR
to digitize these files into text which is suitable for analysis. This
means there are considerable digitization errors especially in the first
half of the dataset.

While there is evidently a growing body of literature on this topic,
there is still no comprehensive database for Australian Hansard based on
XML that spans from 1901 to the present day. Our work serves to bridge
this gap.

\hypertarget{methods}{%
\section{Methods}\label{methods}}

Our database contains one comma-separated value (CSV) file for each
sitting day of the House of Representatives from 02 March 1998 to 08
September 2022. We developed four scripts to produce all of these files,
where each script is designed to parse Hansard documents from a specific
portion of the 1998 to 2022 time frame.

This section will be structured as follows. First, we will provide an
overview of our approach to understanding and parsing an individual
Hansard XML document, which informed all four scripts used to create our
database. This will be supplemented with an excerpt from a Hansard XML
to provide a visual example of its structure. Next we will explain the
specific differences between the four scripts, and outline what
structural changes necessitated their separate development. We will then
provide details on the methodological intricacies of three core
components of Hansard proceedings: Question Time, interjections, and
stage directions. Finally, we will discuss the script we developed to
fill in remaining missing speaker details, which each file in our
database was passed to after being parsed and cleaned.

\hypertarget{sec-overview}{%
\subsection{Overview}\label{sec-overview}}

The approach to parsing contents of an XML document is heavily dependent
on its tree structure. As such, to create this database, we started by
looking at a single Hansard XML transcript from 2019. Doing so enabled
us to identify the various components of interest in the document, and
how each one can be parsed according to its corresponding structural
form. Parsing was performed in \texttt{R} using the \texttt{XML} and
\texttt{xml2} packages (Temple Lang 2022; Wickham, Hester, and Ooms
2021). Focusing on one transcript also allowed us to ensure that all key
components of the transcript were being parsed and captured with as much
detail as possible. The typical form of a Hansard XML transcript is
summarized in the nested list below. Note that this provides a broad
overview of the structure, and does not contain every possible nested
element that may be found in a Hansard XML.

\begin{quote}
\begin{verbatim}
     <hansard>
       1. <session.header>
       2. <chamber.xscript>
          a)  <business.start>
          b)  <debate>
              i.  <debateinfo>
              ii. <debate.text>
              iii. <speech>
              iv. <subdebate.1>
                  (1) <subdebateinfo>
                  (2) <subdebate.text>
                  (3) <speech>
                  (4) <subdebate.2>
                      (a) <subdebateinfo>
                      (b) <subdebate.text>
                      (c) <speech>
      3. <fedchamb.xscript>
      4. <answers.to.questions>
          a)  <question>
          b)  <answer>
\end{verbatim}
\end{quote}

The outer-most node, also known as the parent node, is denoted
\texttt{\textless{}hansard\textgreater{}} and serves as a container for
the entire document. This parent node may have up to four child-nodes,
where the first child-node contains details on the specific sitting day.
Next, \texttt{\textless{}chamber.xscript\textgreater{}} contains all
proceedings of the Chamber,
\texttt{\textless{}fedchamb.xscript\textgreater{}} contains all
proceedings of the Federation Chamber, and
\texttt{\textless{}answers.to.questions\textgreater{}} contains Question
Time proceedings. Note that the Federation Chamber does not meet on
every sitting day, so this child element is not present in every XML
file. The use of separate child nodes allows for the distinction of
proceedings of the Chamber and Federation Chamber. The structure of the
\texttt{\textless{}chamber.xscript\textgreater{}} and
\texttt{\textless{}fedchamb.xscript\textgreater{}} nodes are generally
the same, where the proceeding begins with a business start which is
followed by a series of debates. Debate nodes can contain a
\texttt{\textless{}subdebate.1\textgreater{}} child-node which has a
\texttt{\textless{}subdebate.2\textgreater{}} child-node nested within
it. That said, sometimes \texttt{\textless{}subdebate.2\textgreater{}}
is not nested within \texttt{\textless{}subdebate.1\textgreater{}}. Each
of these three elements (i.e.~\texttt{\textless{}debate\textgreater{}},
\texttt{\textless{}subdebate.1\textgreater{}}, and
\texttt{\textless{}subdebate.2\textgreater{}}) as well as their
respective sub-elements contain important information on the topic of
discussion, who is speaking, and what is being said. The
\texttt{\textless{}speech\textgreater{}} node within each one contains
the bulk of the text associated with that debate or sub-debate. A
typical \texttt{\textless{}speech\textgreater{}} node begins with a
\texttt{\textless{}talk.start\textgreater{}} sub-node, providing
information on the speaker and the time of their first statement.
Unsurprisingly, speeches rarely go uninterrupted in parliamentary debate
settings --- they are often composed of a series of interjections and
continuations. These statements are categorized under different
sub-nodes depending on their nature, such as
\texttt{\textless{}interjection\textgreater{}} or
\texttt{\textless{}continuation\textgreater{}}. The final key component
of Hansard is Question Time, in which questions and answers are
classified as unique elements. More detail on the purpose and processing
of Question Time will follow in Section~\ref{sec-qa}.

Figure~\ref{fig-xml1} provides an example of the beginning of an XML
file for Hansard, which illustrates the structure outlined in the nested
list above. As stated, the XML structure begins with a parent element
\texttt{\textless{}hansard\textgreater{}} (highlighted in blue),
followed by a child element
\texttt{\textless{}session.header\textgreater{}} (highlighted in yellow)
with sub-child elements such as the date and parliament number, which
are all highlighted in pink. Next there is the child element containing
everything that takes place in the Chamber,
\texttt{\textless{}chamber.xscript\textgreater{}}, which is also
highlighted in yellow in Figure~\ref{fig-xml1}. As previously mentioned,
the first sub-node of \texttt{\textless{}chamber.xscript\textgreater{}}
is the business start. The structure of the business start can be seen
between the nodes highlighted in green in Figure~\ref{fig-xml1}, where
the content we parse from the business start is highlighted in orange.

\begin{figure}

{\centering \includegraphics{images/xml1.png}

}

\caption{\label{fig-xml1}Snapshot of the beginning of the XML file for
Hansard on 25 February 2020}

\end{figure}

Evidently, the nature of XML formatting means that different pieces of
information are categorized under a series of uniquely named and nested
nodes. As a result, to parse each piece of information, one must specify
the unique hierarchy of nodes in which it is structured. This is known
as an XPath expression, and tells the parser how to navigate the XML
document to obtain the desired information. For example, the session
header date in Figure~\ref{fig-xml1} can be accessed using the XPath
expression ``\texttt{hansard/session.header/date}''. When specifying an
XPath expression, one can use an ``or'' operator to obtain elements from
multiple node paths at once, in the order that they appear in the
document. We did so throughout our script as we parsed uniquely nested
speech content, as it allows for the correct ordering of elements to be
maintained. We began our first script by parsing all the business start,
speech text, and Question Time contents contained in the XML document,
using these unique XPath expressions to do so. Hansard records contain a
number of statements throughout that were not explicitly spoken in
parliament. For example, the text in Figure~\ref{fig-content_ex} beneath
the bold subtitle ``Report from Federation Chamber'' is not attributed
to a particular member and was not spoken aloud during the parliamentary
proceedings. In contrast, the text which comes afterward was spoken
aloud in parliament by The Speaker at 14:01. Generally, unspoken text is
nested under \texttt{\textless{}debate.text\textgreater{}} or
\texttt{\textless{}subdebate.text\textgreater{}} nodes, while spoken
text is categorized under \texttt{\textless{}speech\textgreater{}}.
Importantly, in our work we focus primarily on parsing spoken speech
content (i.e.~what is nested within
\texttt{\textless{}speech\textgreater{}} nodes), such as what is stated
by The Speaker in Figure~\ref{fig-content_ex}. For this reason, many
Hansard transcriptions which are not spoken text may not be captured in
our database. In future work, capturing this unspoken content in our
database would be a priority.

\begin{figure}

{\centering \includegraphics[width=1.88542in,height=\textheight]{images/content_ex.png}

}

\caption{\label{fig-content_ex}Snapshot from page 7129 of the PDF file
for Hansard on 26 March 2013}

\end{figure}

The next step was to further develop our script to produce tidy data
sets containing all parsed text elements, where each statement is
separated onto its own row with details about the speaker, and rows are
maintained in chronological order. This first involved correcting the
variable classes and adding several indicator variables to differentiate
where statements came from, such as Chamber versus Federation Chamber or
sub-debate 1 versus sub-debate 2. The next key task stemmed from the
fact that the raw text data were not separated by each statement when
parsed. In other words, any interjections, comments made by The Speaker
or Deputy Speaker and continuations within an individual speech were all
parsed together as a single string. As such, the name, name ID,
electorate and party details were only provided for the person whose
turn it was to speak. Splitting up these speeches in a way which would
be generalizable across sitting days required much thought and effort.
Section~\ref{sec-interject} will provide further details on the
intricacies of this task.

Since we are looking at a wide time span of documents, there are many
changes in the way they are formatted over time. These became apparent
as we ran our script on XML files from earlier sitting days. Some
changes are as subtle as a differently named child-node, while others
are as extensive as a completely different nesting structure. Smaller
changes were accounted for as we became aware of them, and embedded into
the code in a way that would not cause issues for parsing more current
Hansards with subtle differences in formatting. However, as mentioned,
more significant changes in the XML structure of Hansard are what
necessitated the development of separate scripts as we worked backwards
in time. Further, while working backward in time, it became clear that
not every sitting day contains every possible XML element. For example,
some days did not have sub-debate 2 content, and some days did not have
a Federation Chamber proceeding. To improve the generalizability of
these scripts, if-else statements were embedded within the code wherever
an error might arise due to a missing element. For example, the entire
Federation Chamber block of code is wrapped in an if-else statement for
each script, so that it only executes if what the code attempts to parse
exists in the file.

Once the script ran without error for a few recent years of Hansard, we
continued to work backwards until extensive changes in tree structure
made our script incompatible with parsing earlier XML files. The
earliest sitting day our first script can successfully parse is 14
August 2012. Before developing new scripts to parse earlier Hansard
documents, we decided to prioritize cleaning and finalizing what we have
been able to parse. As such we continued building our script, fixing any
problems we noticed in the resulting datasets such as excess whitespace
or spacing issues, and splitting up any additional sections of the
parsed text onto separate rows where necessary. Specifically, we added a
section of our script to separate out general stage directions. More
information on this separation will be provided in
Section~\ref{sec-stage}. After completing our first script, it was
formatted as a function which takes a single file name argument and
produces one CSV file containing data on all proceedings from the given
sitting day.

\hypertarget{sec-diff}{%
\subsection{Script Differences}\label{sec-diff}}

As previously mentioned, we developed a total of four scripts to parse
the 1998-2022 time frame of Hansard documents. Two main factors
motivated us to create four scripts as opposed to just one, the first
being structural variation in XML over time, and the second being
improved computational efficiency with separate scripts. While all four
scripts use the same general approach to parsing described in
Section~\ref{sec-overview} and produce the same CSV structure, the first
and second scripts use a different method of data processing than the
third and fourth scripts.

The need for a second script stems from the fact that when established
in 1994, the Federation Chamber was originally named the Main Committee.
The Main Committee was renamed to the Federation Chamber in mid-2012
(House of Representatives 2018, ch.~21). As a result, the child-node
under which Federation Chamber proceedings are nested is named
\texttt{\textless{}maincomm.xscript\textgreater{}} in all XML files
prior to 14 August 2012. Having developed our first script based on
Hansard from recent years, all XPath expressions contain the
\texttt{\textless{}fedchamb.xscript\textgreater{}} specification. To
avoid causing issues in our first script which successfully parses about
10 years of Hansard, we created a second script where we replaced all
occurrences of \texttt{\textless{}fedchamb.xscript\textgreater{}} with
\texttt{\textless{}maincomm.xscript\textgreater{}}. After making this
modification and accounting for other small changes such as time stamp
formatting, this second script successfully parses all Hansard sitting
days from 10 May 2011 to 28 June 2012 (inclusive).

While the modifications needed to develop the second script were quite
straightforward, this was not the case for our next script. The typical
tree structure of Hansard XMLs spanning from 1998 to March 2011 has an
important difference from that of XMLs released after March 2011,
necessitating many changes to be made in our methodology. In XMLs after
March 2011, which our first two scripts successfully parse, the first
two child-nodes of \texttt{\textless{}speech\textgreater{}} are
typically \texttt{\textless{}talk.start\textgreater{}}, and
\texttt{\textless{}talk.text\textgreater{}}. The first child-node
contains data on the person whose turn it is to speak, and the second
contains the entire contents of that speech --- including all
interjections, comments, and continuations. After the
\texttt{\textless{}talk.text\textgreater{}} element closes, there are
typically a series of other child-nodes which provide a skeleton
structure for how the speech proceedings went in chronological order.
For example, if the speech began, was interrupted by one member, and
then continued uninterrupted until the end, there would be one
\texttt{\textless{}interjection\textgreater{}} node and one
\texttt{\textless{}continuation\textgreater{}} node following the
\texttt{\textless{}talk.text\textgreater{}} node. These would contain
details on the person who made each statement such as their party and
electorate.

In contrast, the speech contents in XMLs from 1998 up to and including
24 March 2011 are nested differently --- there is no
\texttt{\textless{}talk.text\textgreater{}} node. Rather than this
single child-node that contains all speech content, statements are
categorized in individual child-nodes. This means that unlike our code
for parsing more current Hansards, we cannot specify a single XPath
expression such as
``\texttt{chamber.xscript//debate//speech/talk.text}'' to extract all
speeches, in their entirety, at once. This difference in nesting
structure made many components of our second script unusable for
processing transcripts preceding 10 May 2011, and required us to change
our data processing approach considerably.

Since the earlier Hansard XMLs do not have a
\texttt{\textless{}talk.text\textgreater{}} node, we found that the most
straightforward way to preserve the ordering of statements and to parse
all speech contents at once was to parse from the
\texttt{\textless{}debate\textgreater{}} element directly. The reason we
did not use its \texttt{\textless{}speech\textgreater{}} child-node is
because every speech has a unique structure of node children, and this
makes it difficult to write code for data cleaning which is
generalizable across all speeches and sitting days. The challenge with
parsing through the \texttt{\textless{}debate\textgreater{}} element is
that every piece of data stored in that element is parsed as a single
string, including all \texttt{\textless{}talk.start\textgreater{}} data,
and all nested sub-debate data. For example, the data shown in
Figure~\ref{fig-patternEX} would be parsed as a single string preceding
the speech content, like so:

\begin{quote}
\texttt{09:31:0010261Costello,\ Peter,\ MPMr\ COSTELLOCT4HigginsLPTreasurer10}
\end{quote}

\begin{figure}

{\centering \includegraphics[width=3.28125in,height=\textheight]{images/patternEX.png}

}

\caption{\label{fig-patternEX}Portion of XML file for Hansard on 12
December 2002}

\end{figure}

This was not isolated to just the beginning of speeches --- details on
individuals interjecting or commenting during speeches were also
captured this way. To separate statements out correctly, we collected
all of these patterns using the
\texttt{\textless{}talk.start\textgreater{}} node, and used them to
split statements wherever one of these patterns was found. After
separating the statements, we were able to remove these patterns from
the body of text. We also used this method of extracting and later
removing unwanted patterns for other pieces of data which did not belong
to the debate proceedings, such as sub-debate titles.

Once we finalized this new method of processing the data, we proceeded
with data cleaning using the same general approach as in the first two
scripts to produce the same structure of CSV output. We then worked
backwards in time and modified the code as needed for generalizability.
Throughout this process we found many transcription errors present in
the XMLs from earlier years. We fixed these manually, deferring to the
official release to ensure the correct information was filled in. Since
there were a number of transcription errors specific to the 2000s, we
chose to create a fourth script for parsing 1998 and 1999. This allowed
us to remove all the code which was needed to resolve specific
transcription errors of the 2000s, to avoid an overly long script and in
turn improving computational efficiency. As such, our fourth script is
essentially the same as the third, with the only difference being that
it has code specific to fixing transcription errors from 1998 and 1999.

\hypertarget{sec-qa}{%
\subsection{Question Time}\label{sec-qa}}

A key characteristic of the Australian parliamentary system is the
ability for the executive government to be held accountable for their
decisions. One core mechanism by which this is achieved is called
Question Time. This is a period of each sitting day in the Chamber where
members of the House can ask ministers two types of questions: questions
in writing which are written in advance, or questions without notice
which are asked verbally in the Chamber and are responded to in real
time (House of Representatives 2021). Questions without notice are
included directly in the \texttt{"chamber.xscript"} child node, with
sub-child nodes called \texttt{"question"} and \texttt{"answer"} to
differentiate the two. Questions in writing, however, are embedded in
their own child node called \texttt{"answers.to.questions"} at the end
of the XML file.

Our approach to parse the \texttt{"chamber.xscript"} speeches used in
all four scripts meant that all questions without notice content was
already parsed in order. For the first two scripts, questions and
answers were already separated onto their own rows. For the third and
fourth scripts, just as we did with the rest of the speech content, we
used those patterns of data preceding the text to separate questions and
answers. Finally, since questions in writing exist in their own child
node we were able to use the same parsing method for all scripts, which
was simply to extract all question and answer elements from the
\texttt{"answers.to.questions"} child node.

We then added binary flags to differentiate between questions and
answers. To do this in the first and second scripts, we separately
re-parsed question and answer content using the XPath expressions
\texttt{"chamber.xscript//question"} and
\texttt{"chamber.xscript//answer"}, added the correct question and
answer flags accordingly, and then added those flags back to the main
dataframe based on exact text matches. For the third and fourth scripts,
we made use of the fact that the patterns preceding text transcribed
under a question node were stored separately from those transcribed
under an answer node. As a result, we could readily use those patterns
to flag questions and answers correctly based on which list of patterns
it belonged to. Sometimes, questions were incorrectly transcribed under
an answer node and vice-versa, in which cases we manually corrected the
question and answer flags using the \texttt{str\_detect} function from
the \texttt{stringr} package. For instance, we check for any statements
flagged as questions which include the phrase ``has provided the
following answer to the honourable member's question'', in which case we
re-code that statement as an answer.

The next step was to merge Question Time contents with all the debate
speech. Doing so with the questions without notice content was
straightforward. As mentioned, our method of parsing meant that
everything was already parsed together in order, so we did not have to
perform any additional merging. For questions in writing, merging this
content was also straightforward due to the fact that it is always at
the very end of Hansard. This means that we could simply bind question
in writing rows to the bottom of the main dataframe. This approach was
used for all four scripts.

\hypertarget{sec-interject}{%
\subsection{Interjections}\label{sec-interject}}

As mentioned, the text was structured and parsed in such a way that
various interjections and comments which happened during a speech were
not separated onto individual rows. This was the case across the entire
time frame of documents. We will first discuss the methodology employed
to split interjections in the first and second scripts, as it informed
our approach for the third and fourth scripts.

Below is an example of part of a speech we would need to split,
extracted from Hansard on 30 November 2021, where Mr.~van Manen is
interrupted by The Speaker who states that the time for members'
statements has concluded.

\begin{quote}
``Mr VAN MANEN (Forde---Chief Government Whip) (13:59): It's a great
pleasure to share with the House that Windaroo Valley State High School
has qualified for the finals of the Australian Space Design Competition,
to begin in January next year. The competition is regarded as the
premier STEM competition for high school students and is recognised by
universities around the country. The students are required to respond to
industry-level engineering and requests for tender for design and---The
SPEAKER: Order! In accordance with standing order 43, the time for
members' statements has concluded.''
\end{quote}

We want each statement on its own row with the correct name, name ID,
electorate and party information on the individual speaking. We
approached this task in a number of steps.

Once all parsed text from the XML was merged into one dataframe called
\texttt{main}, our first step was to add a \texttt{"speech\_no"}
variable. This was done to keep track of which speech each interjection,
comment, or continuation belonged to as we separated these components
onto their own rows.

The next step was to extract all the names and titles preceding these
interjections, comments and continuations. This would enable us to then
separate the speeches in the correct places using these names and titles
in combination with regular expressions, which are patterns of
characters that can be used to search bodies of text. We completed this
extraction process with a few intermediate steps, due to the large
number of name styles and interjection types that had to be accounted
for, each requiring their own unique regular expression format.

As mentioned in Section~\ref{sec-diff}, more recent years of Hansard
XMLs contain a series of child-nodes which exist to capture the
structure of interruptions in that speech. Figure~\ref{fig-xml3}
provides an example of this, where the speech was interrupted by a
comment from The Deputy Speaker, and then the member continued their
speech. Looking at the element names highlighted in blue, these
child-nodes do not contain the actual text for the interjection or
continuation --- this text is embedded within the speech above it.
However, as shown by the content highlighted in pink in
Figure~\ref{fig-xml3}, we were able to extract useful details on the
individual interjecting which we could use later. Making use of this
structure, we extracted names and information of all individuals that
were categorized within the XML as interjections. We stored this as a
dataframe called \texttt{"interject"}.\footnote{We decided not to
  include this data in our final database, as it is all embedded in our
  resulting datasets which have a flag for interjections.}

\begin{figure}

{\centering \includegraphics[width=3.59375in,height=\textheight]{images/xml3.png}

}

\caption{\label{fig-xml3}Snapshot of XML structure with interjection and
continuation from 03 February 2021 Hansard}

\end{figure}

We then created lists using both the \texttt{interject} and
\texttt{main} dataframes to capture all the names of individuals who
spoke that day. We added the names of all Members in a number of unique
formats, due to the frequent variation in how names are transcribed in
Hansard. When a Member interjects or continues a speech, the usual form
of their name is a title followed by their first name or first initial
and/or last name. There is also variation in the capitalization of these
names.\footnote{Sometimes when someone's first name is included, only
  their last name is capitalized, while sometimes their full name is
  capitalized, or other times neither are capitalized.} Another source
of variation is in individuals with more than one first name, as
sometimes only their initial first name is written, while other times
their entire first name is written. Additionally, some surnames have
punctuation, and some surnames have specific capitalization such as
``McCormack'', where even in full capitalization, the first ``c''
remains lower case. This variation demands careful consideration when
writing regular expression patterns to search for in the text. In these
lists we also accounted for any general interjection statements
transcribed that were not attributed to an individual, such as ``An
opposition member interjecting-''.

Having these lists enabled us to extract the names of Members and their
associated titles as they exist in the text, by searching for exact
matches with regular expression patterns. We then used these extracted
names to split all the speeches up, using regular expressions with
lookarounds. A lookaround can be added to a regular expression pattern
to enhance the specificity of matches. These were used to ensure that
the text was not being split in the wrong places, such as places where
Members were simply being named in the statement of another Member.

Once all interjections, comments and continuations were successfully
split onto their own rows using the lists we created, we did one final
check for any additional names that were not captured in these lists. To
do so, we searched for any remaining name matches in speech bodies with
general regular expressions and lookarounds, and separated text using
those matches when found.

We then added an order variable to the dataset based on row number, to
keep track of the order in which everything was said. The next step was
to fill the name, name ID, electorate and party variables with the
correct data for each row. We also wanted to add the gender and unique
identifier for each individual as found in the
\texttt{AustralianPoliticians} package. To do so, we created a lookup
table, which contained the unique incomplete form in which the name was
transcribed, and the corresponding full name, name ID, electorate,
party, gender, and unique ID for that individual.
Figure~\ref{fig-lookup} provides an example of this. We used the main
dataset from the \texttt{AustralianPoliticians} package in the creation
of each lookup table (Alexander and Hodgetts 2021).

\begin{figure}

{\centering \includegraphics[width=4.66667in,height=\textheight]{images/lookup_ex.png}

}

\caption{\label{fig-lookup}First 10 rows of lookup table from 19 October
2017 Hansard processing}

\end{figure}

Next, we merged our main dataframe with the lookup table to replace any
incomplete names with their full names, and to fill in any gaps with
available name ID, electorate, party, gender, and unique ID information.
Finally, we were able to add a flag for interjections. Grouping our data
by the speech number, we defined an interjection as a statement made by
anyone who is not The Speaker, The Deputy Speaker, or the Member whose
turn it was to speak. Figure~\ref{fig-interject} provides an example of
a Federation Chamber proceeding with interjections. Evidently,
statements made by the original speaker Stuart Robert or by The Deputy
Speaker Maria Vamvakinou are not flagged as interjections.

\begin{figure}

{\centering \includegraphics{images/interject_ex.png}

}

\caption{\label{fig-interject}Example of speech with interjections from
21 November 2016 Hansard}

\end{figure}

Having developed a successful methodology for splitting interjections,
we used this to inform our general approach in the third and fourth
scripts. However, the difference in data cleaning used in these scripts
necessitated some departure from the original methodology. As discussed
in Section~\ref{sec-diff}, we used string patterns extracted from
\texttt{\textless{}talk.start\textgreater{}} nodes to separate speeches.
As evident in Figure~\ref{fig-xml3},
\texttt{\textless{}talk.start\textgreater{}} nodes are nested within
\texttt{\textless{}interjection\textgreater{}} nodes, meaning that the
patterns of data from interjection statements were separated out in the
process. In other words, our approach to data cleaning in the third and
fourth scripts separated out interjections in the process. This meant
that we did not need to create lists of names and titles to search for
in the text as we did before. However, we used the same list of general
interjection statements to separate on that was used in the first two
scripts. We then did an additional check for statements that may have
not been separated due to how they were embedded in the XML, and
separated those out where needed.\footnote{While most statements were
  categorized in their own child-node and hence captured through
  pattern-based separation, some were not individually categorized, and
  had to be split manually in this step.}

We then proceeded to clean up speeches and fill in correct speaker
details. While we used the same lookup table approach as before, we did
so in combination with another means of filling in speaker details. The
patterns parsed from \texttt{\textless{}talk.start\textgreater{}} nodes
contain important data on the speaker of each statement. As such, we
could extract those data associated with each pattern by parsing one
element inward, using the XPath expression
\texttt{\textless{}talk.start/talker\textgreater{}}. We created a
pattern lookup table with these data, and merged it with the main
Hansard dataframe by the first pattern detected in each statement.
Figure~\ref{fig-patterns} provides an example of that lookup table. This
approach enabled us to fill in missing data on each speaker using data
extracted directly from the XML. Finally, we then used the
\texttt{AustralianPoliticians} dataset to fill in other missing data,
and flagged for interjections in the same manner as before.

\begin{figure}

{\centering \includegraphics{images/pattern_lookup.png}

}

\caption{\label{fig-patterns}10 rows of pattern lookup table from 12
December 2012 Hansard processing}

\end{figure}

\hypertarget{sec-stage}{%
\subsection{Stage Directions}\label{sec-stage}}

When building our first scripts, one of the final components needed was
to separate general stage directions out from statements made by
members. Stage directions are general statements included in the
transcript to document happenings in parliament. Examples of stage
directions are ``Bill read a second time'', ``Question agreed to'' or
``Debate adjourned''. It was unclear to us from the XML and PDF who
exactly these statements were attributed to. For further clarification,
we watched portions of the video recording for some sitting days, and
noticed that where these statements are documented in Hansard, they are
not explicitly stated in parliament. For example, when The Deputy
Speaker says ``The question is that the bill be now read a second
time'', members of the House take a vote, and if the majority is in
favour, they proceed reading the bill the second time. This vote and
second reading is not explicitly transcribed, rather what is written is:
``Question agreed to. Bill read a second time''. For this reason, we
filled the name variable for these statements with ``stage direction''.
Note that stage directions were not flagged as interjections. These
stage directions are not defined differently from the regular debate
speech in the XML, meaning we had to manually create a list of stage
directions to separate out of the speeches. We have built this list of
stage directions as we worked backwards in parsing Hansard, and took the
same approach across all four scripts.

\hypertarget{filling-missing-details}{%
\subsection{Filling Missing Details}\label{filling-missing-details}}

While we did our best to maximize the completeness of the files in our
database as they were processed in the initial four scripts, there were
still a number of rows in which details on the person speaking were
missing, or the name transcribed for that individual was in a short form
(i.e.~``Mr Abbott'' instead of ``Abbott, Tony, MP''). This was a
particularly frequent occurrence for sitting days where a member spoke
whose surname was shared by any other past or present member of
parliament, and automated filling of their details using data from the
\texttt{AustralianPoliticians} package was avoided to prevent any
incorrect detail attribution. In an effort to improve as many of these
as possible, we developed a script which identifies short-form names
belonging to people with common surnames in each CSV, looks for the full
version of that individuals name if available in that same CSV file, and
replaces the short-form name with the full name, and fills the rest of
the speaker details in accordingly with data from the
\texttt{AustralianPoliticians} package. This script does the same for
anyone who does have a unique surname but is still missing the full name
form or any gender, unique ID, name ID, party or electorate details.
Each file in our database passed through this script after being
created, to ensure it is as complete as possible.

\hypertarget{data-records}{%
\section{Data Records}\label{data-records}}

Our database contains 1517 CSV files, covering all sitting days of the
House of Representatives from 02 March 1998 to 08 September 2022. All
data records are available on the general-purpose repository Zenodo, at
https://doi.org/10.5281/zenodo.7336076 (Katz and Alexander 2022). For
each CSV file, each row contains an individual statement, with details
on the individual speaking. Of course, for general statements
transcribed as made by ``Honourable members'' for example, these
variables cannot be specified. Table~\ref{tbl-vars} provides an overview
of each variable found in the database.

\hypertarget{tbl-vars}{}
\begin{table}[H]
\caption{\label{tbl-vars}Summary and description of variables in our database }\tabularnewline

\centering
\begin{tabular}{ll}
\toprule
Variable & Description\\
\midrule
name & Name of speaker\\
order & Row number\\
speech\_no & Speech number\\
page.no & Page number statement can be found on in official Hansard\\
time.stamp & Time of statement\\
\addlinespace
name.id & Unique member identification code, based on the Parliamentary Handbook\\
electorate & Speaking member's electorate\\
party & Speaking member's party\\
in.gov & Flag for in government (1 if in government, 0 otherwise)\\
first.speech & Flag for first speech (1 if first speech, 0 otherwise)\\
\addlinespace
body & Statement text\\
fedchamb\_flag & Flag for Federation Chamber (1 if Federation Chamber, 0 if Chamber)\\
sub1\_flag & Flag for sub-debate 1 contents (1 if sub-debate 1, 0 otherwise)\\
sub2\_flag & Flag for sub-debate 2 contents (1 if sub-debate 2, 0 otherwise)\\
question & Flag for question (1 if question, 0 otherwise)\\
\addlinespace
answer & Flag for answer (1 if answer, 0 otherwise)\\
q\_in\_writing & Flag for question in writing (1 if question in writing, 0 otherwise)\\
div\_flag & Flag for division (1 if division, 0 otherwise)\\
gender & Gender of speaker\\
uniqueID & Unique identifier of speaker\\
\addlinespace
interject & Flag for interjection (1 if statement is an interjection, 0 otherwise)\\
\bottomrule
\end{tabular}
\end{table}

The \texttt{name}, \texttt{page.no}, \texttt{time.stamp},
\texttt{name.id}, \texttt{electorate}, \texttt{party}, \texttt{in.gov},
\texttt{first,speech}, and \texttt{body} variables all came directly
from the XML contents. In addition to these variables, we added a number
of flags to enable easy filtering of statements. For example, adding the
\texttt{fedchamb\_flag} provides a clear distinction between the
proceedings of the Chamber with those of the Federation Chamber. As
well, the \texttt{sub1\_flag} and \texttt{sub2\_flag} variables allow us
to keep track of where various statements are being parsed from in the
XML document. The \texttt{question}, \texttt{answer}, and
\texttt{q\_in\_writing} flags were added to identify statements
belonging to Question Time, and the nature of these statements. We also
flagged for interjections (\texttt{interject}), and the
\texttt{div\_flag} variable was added to flag when a division is called
for. The \texttt{gender} and \texttt{uniqueID} variables were added
based on the main dataset from the \texttt{AustralianPoliticians}
package. Details on the usage of \texttt{uniqueID} will be provided in
Section~\ref{sec-usage}. Further, the \texttt{speech\_no} variable
allows us to keep track of the speech number that each statement and
interjection belongs to. Having the speech number variable offers an
easy way to group statements by speech or isolate specific speeches of
interest. Finally, the \texttt{order} variable was added to maintain the
order of proceedings.

\hypertarget{technical-validation}{%
\section{Technical Validation}\label{technical-validation}}

We developed a script to perform seven automated tests on each file in
our database, in an effort to enhance its quality and consistency. Our
first test validates that the date specified in each file name matches
the date specified in its corresponding XML session header.\footnote{As
  seen in Figure~\ref{fig-xml1}, the first child node of the
  \texttt{\textless{}session.header\textgreater{}} element is the date.}
Every file passed this test, and we detected one discrepancy in an XML
file from 03 June 2009, where its session header contained the wrong
date. We validated that our file name and date was correct by checking
the official PDF release from that sitting day.

The second test is designed to test for duplication errors in the data,
by checking whether two lines that immediately follow each other have
the same body (i.e.~spoken content). This test detected 129 dates on
which duplicate statements were made, and one immediately follows the
other. Note that this test does not account for who is making each
statement, meaning one member repeating the words of another member
would be picked up in this test as well. We checked a sample of 40\% of
these duplicates, and manually validated that they are all repeated
statements that do exist and are transcribed closely together in that
day's XML file, and by our method should be parsed such that one of
these statements is immediately followed by the other.

When a member runs out of allotted time for their speech, Hansard
editors transcribe ``(Time expired)'' after their final word. As a means
of checking that we have separated speeches out correctly, our third
test checks that when the phrase ``(Time expired)'' exists in a body of
text, it exists at the very end. When this is not the case, we know that
we have missed the separation of the next statement onto its own row,
and could fix this accordingly.

The remaining tests focus on the members present on each sitting day.
Our fourth test checks that there is one unique party and electorate
attributed to each individual on each sitting day. As we parsed Hansard
further back in time, we found a number of cases where an individual was
associated with the wrong electorate or party due to transcription
errors. When we found these data errors we corrected them based on the
official release. This test provides us with an automated way to catch
these errors and correct them at scale.

Next, we test that the unique name identification code attributed to
each individual is found in the Australian Parliamentary Handbook. We do
so using the \texttt{ausPH} package. This test serves as another means
to correct for transcription errors, this time in the case of name IDs.
We found and corrected for a number of common name ID transcription
errors detected by this test, such as a capital letter ``O'' in place of
a zero.

Our sixth test checks that on any given sitting day, the individuals
identified are alive. To do so, we utilized the main dataset from the
\texttt{AustralianPoliticians} package which contains the birth and
where applicable death dates for every politician. This test confirmed
that all members who are detected to speak on each sitting day are not
deceased.

Finally, our seventh test validates that all individuals speaking are
active members of parliament on that particular day. We use the
\texttt{mps} dataset from the \texttt{AustralianPoliticians} package
which has the dates of each Member's duration in parliament. Using these
dates, we check that each person speaking on each sitting day is in fact
a Member of Parliament on that day.

\hypertarget{sec-usage}{%
\section{Usage Notes}\label{sec-usage}}

To enhance the usability of our database, we added a \texttt{uniqueID}
variable to each CSV file. This serves as a unique identifier for each
speaking Member, and comes from the \texttt{uniqueID} variable present
within data from both the \texttt{AustralianPoliticians} and
\texttt{AustralianElections} R packages (Alexander and Hodgetts 2021;
Alexander 2019). By including this variable, one can readily integrate
our data records with those available in these two packages.

Further, the \texttt{name.id} variable found in each CSV is another
unique identifier for each Member of Parliament. This variable was
parsed directly from the Hansard XML files, and can be found in the
Australian Parliamentary Handbook. As such, our data records can be
integrated with those from the \texttt{ausPH} package which provides
datasets for contents of the Australian Parliamentary Handbook (Leslie
2021). This will allow for convenient extraction of further details on
each Member of Parliament in a tidy, ready to analyze format.

\hypertarget{code-availability}{%
\section{Code Availability}\label{code-availability}}

The code written to build this database is available on the GitHub
repository associated with this paper:
https://github.com/lindsaykatz/hansard-proj. All scripts were created
using R software (R Core Team 2022). The core packages used to develop
these scripts are: the \texttt{XML} package by Temple Lang (2022), the
\texttt{xml2} package by Wickham, Hester, and Ooms (2021), the
\texttt{tidyverse} R packages by Wickham et al. (2019), the
\texttt{AustralianPoliticians} package by Alexander and Hodgetts (2021),
and the \texttt{ausPH} package by Leslie (2021). \texttt{XML} and
\texttt{xml2} were used for parsing the XML documents,
\texttt{AustralianPoliticians} and \texttt{ausPH} were used for cleaning
up and filling in member details in the datasets, and \texttt{tidyverse}
packages were used in all steps, for tidy wrangling of data.

\hypertarget{acknowledgements}{%
\section{Acknowledgements}\label{acknowledgements}}

We thank Kristine Villaluna, Monica Alexander, and Jack Stephenson for
helpful comments.

\hypertarget{author-contributions}{%
\section{Author Contributions}\label{author-contributions}}

LK developed and implemented the code to obtain, create, and test, the
datasets and wrote the first draft of the paper. RA conceptualized and
designed the study, and contributed writing. Both authors approved the
final version.

\hypertarget{competing-interests}{%
\section{Competing Interests}\label{competing-interests}}

The authors declare no competing interests.

\hypertarget{references}{%
\section*{References}\label{references}}
\addcontentsline{toc}{section}{References}

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\vadjust pre{\hypertarget{ref-auselect}{}}%
Alexander, Rohan. 2019. \emph{AustralianElections: Datasets on
Australian Elections}.
\url{https://github.com/RohanAlexander/AustralianElections}.

\leavevmode\vadjust pre{\hypertarget{ref-alexander2021}{}}%
Alexander, Rohan, and Monica Alexander. 2021. {``The Increased Effect of
Elections and Changing Prime Ministers on Topics Discussed in the
Australian Federal Parliament Between 1901 and 2018.''} \emph{arXiv
Preprint arXiv:2111.09299}.

\leavevmode\vadjust pre{\hypertarget{ref-auspol}{}}%
Alexander, Rohan, and Paul A. Hodgetts. 2021.
\emph{AustralianPoliticians: Provides Datasets about Australian
Politicians}.
\url{https://CRAN.R-project.org/package=AustralianPoliticians}.

\leavevmode\vadjust pre{\hypertarget{ref-lipad}{}}%
Beelen, Kaspar, Timothy Alberdingk Thijm, Christopher Cochrane, Kees
Halvemaan, Graeme Hirst, Michael Kimmins, Sander Lijbrink, et al. 2017.
{``Digitization of the Canadian Parliamentary Debates.''} \emph{Canadian
Journal of Political Science/Revue Canadienne de Science Politique} 50
(3): 849--64.

\leavevmode\vadjust pre{\hypertarget{ref-fraussen2018}{}}%
Fraussen, Bert, Timothy Graham, and Darren R Halpin. 2018. {``Assessing
the Prominence of Interest Groups in Parliament: A Supervised Machine
Learning Approach.''} \emph{The Journal of Legislative Studies} 24 (4):
450--74.

\leavevmode\vadjust pre{\hypertarget{ref-australia2018house}{}}%
House of Representatives. 2018. Edited by D. R. Elder and P. E. Fowler.
\emph{{House of Representatives Practice}}. 7th ed. {Australian
Government - Department of the House of Representatives}.

\leavevmode\vadjust pre{\hypertarget{ref-house2021}{}}%
---------. 2021. \emph{{A window on the house: Practices and procedures
relating to question time}}. {Parliament of Australia}.

\leavevmode\vadjust pre{\hypertarget{ref-katz_2022}{}}%
Katz, Lindsay, and Rohan Alexander. 2022. {``{A new, comprehensive
database of all proceedings of the Australian Parliamentary Debates}.''}
Zenodo. \url{https://doi.org/10.5281/zenodo.7336076}.

\leavevmode\vadjust pre{\hypertarget{ref-Leslie_2021_HoR_Aus}{}}%
Leslie, Patrick. 2021. \emph{Members of the Australian House of
Representatives (1945-2019)}.
\url{https://github.com/palesl/AustralianHouseOfRepresentatives}.

\leavevmode\vadjust pre{\hypertarget{ref-r_software}{}}%
R Core Team. 2022. \emph{R: A Language and Environment for Statistical
Computing}. Vienna, Austria: R Foundation for Statistical Computing.
\url{https://www.R-project.org/}.

\leavevmode\vadjust pre{\hypertarget{ref-rasiah2010framework}{}}%
Rasiah, Parameswary. 2010. {``A Framework for the Systematic Analysis of
Evasion in Parliamentary Discourse.''} \emph{Journal of Pragmatics} 42
(3): 664--80.

\leavevmode\vadjust pre{\hypertarget{ref-salisbury2011}{}}%
Salisbury, Christopher. 2011. {``'Mr Speaker, i Withdraw...': Standards
of (Mis) Behaviour in the Queensland, Western Australian and
Commonwealth Parliaments Compared via Online Hansard.''}
\emph{Australasian Parliamentary Review} 26 (1): 166--77.

\leavevmode\vadjust pre{\hypertarget{ref-sherratt}{}}%
Sherratt, Tim. 2016. {``Documentation: Historic Hansard.''}
\emph{Historic Hansard}.
\url{http://timsherratt.org/digital-heritage-handbook/docs/historic-hansard/\#}.

\leavevmode\vadjust pre{\hypertarget{ref-XML_pkg}{}}%
Temple Lang, Duncan. 2022. \emph{XML: Tools for Parsing and Generating
XML Within r and s-Plus}. \url{https://CRAN.R-project.org/package=XML}.

\leavevmode\vadjust pre{\hypertarget{ref-vice2017history}{}}%
Vice, John, and Stephen Farrell. 2017. \emph{The History of Hansard}.
House of Lords Library; House of Lords Hansard.

\leavevmode\vadjust pre{\hypertarget{ref-tidyverse}{}}%
Wickham, Hadley, Mara Averick, Jennifer Bryan, Winston Chang, Lucy
D'Agostino McGowan, Romain François, Garrett Grolemund, et al. 2019.
{``Welcome to the {tidyverse}.''} \emph{Journal of Open Source Software}
4 (43): 1686. \url{https://doi.org/10.21105/joss.01686}.

\leavevmode\vadjust pre{\hypertarget{ref-xml2_pkg}{}}%
Wickham, Hadley, Jim Hester, and Jeroen Ooms. 2021. \emph{Xml2: Parse
XML}. \url{https://CRAN.R-project.org/package=xml2}.

\end{CSLReferences}

\end{document}