### 2023-04-05 paper/paper-nature_format.tex
more paper edits

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  letterpaper,
  DIV=11,
  numbers=noendperiod]{scrartcl}

\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{5}
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\KOMAoption{captions}{tableheading}
\makeatletter
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[many]{tcolorbox}}
\makeatother
\makeatletter
\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
\makeatother
\makeatletter
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Digitization of the Australian Parliamentary Debates, 1998-2022},
  pdfauthor={Lindsay Katz; Rohan Alexander},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\title{Digitization of the Australian Parliamentary Debates, 1998-2022}
\author{Lindsay Katz\footnote{University of Toronto,
  lindsay.katz@mail.utoronto.ca} \and Rohan Alexander\footnote{University
  of Toronto, rohan.alexander@utoronto.ca}}
\date{April 5, 2023}

\begin{document}
\maketitle
\begin{abstract}
Public knowledge of what is said in parliament is a tenet of democracy,
and a critical resource for political science research. In Australia,
following the British tradition, the written record of what is said in
parliament is known as Hansard. While the Australian Hansard has always
been publicly available, it has been difficult to use for the purpose of
large-scale macro- and micro-level text analysis because it has only
been available as PDFs or XMLs. Following the lead of the Linked
Parliamentary Data project which achieved this for Canada, we provide a
new, comprehensive, high-quality, rectangular database that captures
proceedings of the Australian parliamentary debates from 1998 to 2022.
The database is publicly available and can be linked to other datasets
such as election results. The creation and accessibility of this
database enables the exploration of new questions and serves as a
valuable resource for both researchers and policymakers.
\end{abstract}
\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[enhanced, borderline west={3pt}{0pt}{shadecolor}, breakable, interior hidden, boxrule=0pt, sharp corners, frame hidden]}{\end{tcolorbox}}\fi

\hypertarget{sec-intro}{%
\section{Background \& Summary}\label{sec-intro}}

The official written record of parliamentary debates, formally known as
Hansard, plays a fundamental role in capturing the history of political
proceedings and facilitating the exploration of valuable research
questions. Originating in the British parliament, the production of
Hansard became tradition in many Commonwealth countries such as Canada,
the United Kingdom, and Australia (Vice and Farrell 2017). Given the
content and magnitude of these records, they have significance,
particularly in the context of political science research. In the case
of Canada, the Hansard has been digitized for 1901 to 2019 (Beelen et
al. 2017). Having a digitized version of Hansard enables researchers to
conduct text analysis and statistical modelling. Following the lead of
that project, in this paper we introduce a similar database for
Australia. This is composed of individual datasets for each sitting day
in the House of Representatives from 1998 to 2022, containing details on
everything said in parliament in a form that can be readily used by
researchers. With the development of tools for large-scale text
analysis, this database will serve as a resource for understanding
political behaviour in Australia over time.

The Australian House of Representatives performs a number of crucial
governmental functions, such as creating new laws and overseeing
government expenditure (House of Representatives 2018, ch.~1). The House
operates under a parallel chamber setup, meaning there are two debate
venues where proceedings take place: the Chamber, and the Federation
Chamber. Sittings of the House follow a predefined order of business,
regulated by procedural rules called standing orders (House of
Representatives 2018, ch.~8). A typical sitting day in the Chamber has a
number of scheduled proceedings including debates on government
business, 90 second member statements, and Question Time (House of
Representatives 2018, ch.~8). The Federation Chamber was created in 1994
as a subordinate debate venue of the Chamber. This allows for better
time management of House business as its proceedings occur
simultaneously with those of the Chamber (House of Representatives 2018,
ch.~21). Sittings in the Federation Chamber are different to those of
the Chamber in terms of their order of business and scope of discussion.
Business matters discussed in the Federation Chamber are limited largely
to intermediate stages of bill development, and the business of private
Members (House of Representatives 2018, ch.~21). It is the recording and
compilation of these proceedings on which Hansard is based, and it is
essentially, but not entirely, verbatim.

A week or so after each sitting day, a transcript is available for
download from the official Parliament of Australia website in both PDF
and extensible markup language (XML) form. The PDF is the official
release. The PDF imposes formatting designed for humans to read with
ease, whereas XML is designed for consistency and machine legibility.
The nature of XML enables us to more easily use code to manipulate these
records at scale, motivating our choice to develop our database solely
using the XML formatted files. In cases where we were unsure on how to
proceed with processing the XML, we defer first to the PDF, and then to
the video recording of the proceeding, if available.

At present, the Hansard format that is available on the Parliament of
Australia website is not accessible for large scale analysis. To this
point, various researchers have had to create their own databases of
usable, complete data based on content from the Australian Parliament
website. For instance, Sherratt (2016) created an online, easy to read
database of Hansard from 1901 to 1980 using the XML files. These data
can be navigated by year, parliament, people, and bills (Sherratt 2016).
To make the Australian Parliamentary Handbook more accessible, Leslie
(2021) has released an \texttt{R} package which includes data on all
Members of the Australian House of Representatives from 1945 to 2019.
Further, Alexander and Hodgetts (2021) created the
\texttt{AustralianPoliticians} \texttt{R} package, which contains
several datasets related to the political and biographical information
of Australian federal politicians who were active between 1901 and 2021.

Many papers exist which use components of Australian Hansard to explore
various research topics. For example, Salisbury (2011) used online
Hansard records to investigate the occurrences of unparliamentary
comments by members of the House, where the Speaker tells that member to
withdraw their remark. Rasiah (2010) worked with Question Time data from
Hansard transcripts during February and March of 2003, to investigate
resistance of politicians in answering questions about Iraq. Fraussen,
Graham, and Halpin (2018) use Hansard to quantify political prominence
by investigating strategic mentions of interest groups by elected
officials. Finally, Alexander and Alexander (2021) construct a dataset
of the Australian Hansard, along with an analysis of the effect of
elections and changes in Prime Ministers upon topics mentioned in
parliament. Alexander and Alexander (2021) created this database with
the static PDF versions of Hansard, using OCR to digitize these files
into text which is suitable for analysis. This means there are
considerable digitization errors especially in the first half of the
dataset.

While there is evidently a growing body of literature on this topic,
there is still no comprehensive database for Australian Hansard based on
XML that spans from 1901 to the present day. Our work begins to bridge
this gap.

\hypertarget{methods}{%
\section{Methods}\label{methods}}

Our database contains one comma-separated value (CSV) and one parquet
file for each sitting day of the House of Representatives from 02 March
1998 to 08 September 2022. We developed four scripts to produce these
files. Each script parses Hansard documents from a specific portion of
the 1998 to 2022 time frame.

This section is structured as follows. First, we provide an overview of
our approach to understanding and parsing an individual Hansard XML
document, which informed the scripts used to create our database. This
will be supplemented with an excerpt from a Hansard XML to provide a
visual example of its structure. Next we will explain the specific
differences between the scripts, and outline what structural changes
necessitated their separate development. We then provide details on the
methodological intricacies of three core components of Hansard
proceedings: Question Time, interjections, and stage directions.
Finally, we discuss the script we developed to fill in remaining missing
details on the member of parliament speaking, which each file in our
database was passed to after being parsed and cleaned.

\hypertarget{sec-overview}{%
\subsection{Overview}\label{sec-overview}}

The approach to parsing contents of an XML document depends on its tree
structure. As such, to create this database, we started by looking at a
single Hansard XML transcript from 2019. Doing so enabled us to identify
the various components of interest in the document, and how each one can
be parsed according to its corresponding structural form. Parsing was
performed in \texttt{R} using the \texttt{XML} and \texttt{xml2}
packages (Temple Lang 2022; Wickham, Hester, and Ooms 2021). Focusing on
one transcript also allowed us to ensure that all key components of the
transcript were parsed and captured in as much detail as possible. The
typical form of a Hansard XML transcript is summarized in the nested
list below. This provides an overview, but does not contain every
possible nested element that may be found in a Hansard XML.

\begin{quote}
\begin{verbatim}
     <hansard>
       1. <session.header>
       2. <chamber.xscript>
          a)  <business.start>
          b)  <debate>
              i.  <debateinfo>
              ii. <debate.text>
              iii. <speech>
              iv. <subdebate.1>
                  (1) <subdebateinfo>
                  (2) <subdebate.text>
                  (3) <speech>
                  (4) <subdebate.2>
                      (a) <subdebateinfo>
                      (b) <subdebate.text>
                      (c) <speech>
      3. <fedchamb.xscript>
      4. <answers.to.questions>
          a)  <question>
          b)  <answer>
\end{verbatim}
\end{quote}

The outer-most node, also known as the parent node, is denoted
\texttt{\textless{}hansard\textgreater{}} and serves as a container for
the entire document. This parent node may have up to four child-nodes,
where the first child-node contains details on the specific sitting day.
Next, \texttt{\textless{}chamber.xscript\textgreater{}} contains all
proceedings of the Chamber,
\texttt{\textless{}fedchamb.xscript\textgreater{}} contains all
proceedings of the Federation Chamber, and
\texttt{\textless{}answers.to.questions\textgreater{}} contains Question
Time proceedings. The Federation Chamber does not meet on every sitting
day, so this child element is not present in every XML file. The use of
separate child nodes allows for the distinction of proceedings between
the Chamber and Federation Chamber. The structure of the
\texttt{\textless{}chamber.xscript\textgreater{}} and
\texttt{\textless{}fedchamb.xscript\textgreater{}} nodes are generally
the same, where the proceeding begins with
\texttt{\textless{}business.start\textgreater{}} which is followed by a
series of debates. Debate nodes can contain a
\texttt{\textless{}subdebate.1\textgreater{}} child-node which has a
\texttt{\textless{}subdebate.2\textgreater{}} child-node nested within
it. That said, sometimes \texttt{\textless{}subdebate.2\textgreater{}}
is not nested within \texttt{\textless{}subdebate.1\textgreater{}}. Each
of these three elements (i.e.~\texttt{\textless{}debate\textgreater{}},
\texttt{\textless{}subdebate.1\textgreater{}}, and
\texttt{\textless{}subdebate.2\textgreater{}}) as well as their
respective sub-elements contain important information on the topic of
discussion, who is speaking, and what is being said. The
\texttt{\textless{}speech\textgreater{}} node within each one contains
the bulk of the text associated with that debate or sub-debate. A
typical \texttt{\textless{}speech\textgreater{}} node begins with a
\texttt{\textless{}talk.start\textgreater{}} sub-node, providing
information on the member of parliament whose turn it is to speak and
the time of their first statement. Unsurprisingly, speeches rarely go
uninterrupted in parliamentary debate settings --- they are often
composed of a series of interjections and continuations. These
statements are categorized under different sub-nodes depending on their
nature, such as \texttt{\textless{}interjection\textgreater{}} or
\texttt{\textless{}continuation\textgreater{}}. The final key component
of Hansard is Question Time, in which questions and answers are
classified as unique elements. More detail on the purpose and processing
of Question Time will follow in Section~\ref{sec-qa}.

Figure~\ref{fig-xml1} provides an example of the beginning of an XML
file for Hansard, which illustrates the structure outlined in the nested
list above. As stated, the XML structure begins with a parent element
\texttt{\textless{}hansard\textgreater{}} (highlighted in blue),
followed by a child element
\texttt{\textless{}session.header\textgreater{}} (highlighted in yellow)
with sub-child elements such as the date and parliament number, which
are all highlighted in pink. Next, there is the child element containing
everything that takes place in the Chamber,
\texttt{\textless{}chamber.xscript\textgreater{}}, which is also
highlighted in yellow in Figure~\ref{fig-xml1}. As previously mentioned,
the first sub-node of \texttt{\textless{}chamber.xscript\textgreater{}}
is \texttt{\textless{}business.start\textgreater{}}. The structure of
this can be seen between the nodes highlighted in green in
Figure~\ref{fig-xml1}, where the content we parse from the business
start is highlighted in orange.

\begin{figure}

{\centering \includegraphics{images/xml1.png}

}

\caption{\label{fig-xml1}Snapshot of the beginning of the XML file for
Hansard on 25 February 2020}

\end{figure}

Evidently, the nature of XML formatting means that different pieces of
information are categorized under a series of uniquely named and nested
nodes. As a result, to parse each piece of information, one must specify
the unique hierarchy of the nodes in which it is structured. This is
known as an XPath expression, and tells the parser how to navigate the
XML document to obtain the desired information. For example, the session
header date in Figure~\ref{fig-xml1} can be accessed using the XPath
expression ``\texttt{hansard/session.header/date}''. When specifying an
XPath expression, one can use an ``or'' operator to obtain elements from
multiple node paths at once, in the order that they appear in the
document. We did so throughout our script as we parsed uniquely nested
speech content. This allows the correct ordering of elements to be
maintained. We began our first script by parsing all the business start,
speech text, and Question Time contents contained in the XML document,
using these unique XPath expressions to do so.

The next step was to further develop our script to produce tidy data
sets, as defined and introduced by Wickham (2014). These contain all
parsed text elements, where each statement is separated onto its own row
with details about the member of parliament who is speaking, and rows
are maintained in chronological order. This first involved correcting
the variable classes and adding several indicator variables to
differentiate where statements came from, such as Chamber versus
Federation Chamber or \texttt{\textless{}subdebate.1\textgreater{}}
versus \texttt{\textless{}subdebate.2\textgreater{}}. The next key task
stemmed from the fact that the raw text data were not separated by each
statement when parsed. In other words, any interjections, comments made
by the Speaker or Deputy Speaker and continuations within an individual
speech were all parsed together as a single string. As such, the name,
name ID, electorate and party details were only provided for the person
whose turn it was to speak. There were many intricacies in the task of
splitting these speeches in a way that would be generalizable across
sitting days. Details on these are provided in
Section~\ref{sec-interject}.

Since we are looking at a wide time span of documents, there are many
changes in the way they are formatted. These became apparent as we ran
our script on XML files from earlier sitting days. Some changes are as
subtle as a differently named child-node, while others are as extensive
as a completely different nesting structure. Smaller changes were
accounted for as we became aware of them, and embedded into the code in
a way that would not cause issues for parsing more current Hansards with
subtle differences in formatting. However, as mentioned, more
significant changes in the XML structure of Hansard are what
necessitated the development of separate scripts as we worked backwards.
Further not every sitting day contains every possible XML element. For
example, some days did not have
\texttt{\textless{}subdebate.2\textgreater{}} content, and some days did
not have a Federation Chamber proceeding. To improve the
generalizability of these scripts, if-else statements were embedded
within the code wherever an error might arise due to a missing element.
For example, the entire Federation Chamber block of code is wrapped in
an if-else statement for each script, so that it only executes if what
the code attempts to parse exists in the file.

Once the script ran without error for a few recent years of Hansard, we
continued to work backwards until extensive changes in tree structure
made our script incompatible with parsing earlier XML files. The
earliest sitting day this first script can successfully parse is 14
August 2012. Before developing new scripts to parse earlier Hansard
documents, we prioritized cleaning and finalizing what we had been able
to parse. As such we continued building our script, fixing any problems
we noticed in the resulting datasets such as excess whitespace or
spacing issues, and splitting up any additional sections of the parsed
text onto separate rows where necessary. Specifically, we added a
section of our script to separate out general stage directions. More
information on this separation will be provided in
Section~\ref{sec-stage}. After completing our first script, it was
formatted as a function which takes a single file name argument and
produces one CSV file containing data on all proceedings from the given
sitting day.

\hypertarget{sec-diff}{%
\subsection{Script Differences}\label{sec-diff}}

As mentioned, we developed a total of four scripts to parse the
1998-2022 time frame of Hansard documents. Two main factors motivated us
to create four scripts as opposed to just one, the first being
structural variation in XML over time, and the second being improved
computational efficiency with separate scripts. While all four scripts
use the same general approach to parsing described in
Section~\ref{sec-overview} and produce the same CSV structure, the first
and second scripts use a different method of data processing than the
third and fourth scripts.

The need for a second script stems from the fact that when established
in 1994, the Federation Chamber was originally named the Main Committee.
The Main Committee was renamed to the Federation Chamber in mid-2012
(House of Representatives 2018, ch.~21). As a result, the child-node
under which Federation Chamber proceedings are nested is named
\texttt{\textless{}maincomm.xscript\textgreater{}} in all XML files
prior to 14 August 2012. Having developed our first script based on
Hansard from recent years, all XPath expressions contain the
\texttt{\textless{}fedchamb.xscript\textgreater{}} specification. To
avoid causing issues in our first script which successfully parses about
10 years of Hansard, we created a second script where we replaced all
occurrences of \texttt{\textless{}fedchamb.xscript\textgreater{}} with
\texttt{\textless{}maincomm.xscript\textgreater{}}. After making this
modification and accounting for other small changes such as time stamp
formatting, this second script successfully parses all Hansard sitting
days from 10 May 2011 to 28 June 2012 (inclusive).

While the modifications needed to develop the second script were
straightforward, this was not the case for our next script. The typical
tree structure of Hansard XMLs spanning from 1998 to March 2011 has an
important difference from that of XMLs released after March 2011,
necessitating many changes to be made in our methodology. In XMLs after
March 2011, which our first two scripts successfully parse, the first
two child-nodes of \texttt{\textless{}speech\textgreater{}} are
typically \texttt{\textless{}talk.start\textgreater{}}, and
\texttt{\textless{}talk.text\textgreater{}}. The first child-node
contains data on the person whose turn it is to speak, and the second
contains the entire contents of that speech --- including all
interjections, comments, and continuations. After the
\texttt{\textless{}talk.text\textgreater{}} element closes, there are
typically a series of other child-nodes which provide a skeleton
structure for how the speech proceedings went in chronological order.
For example, if the speech began, was interrupted by one member, and
then continued uninterrupted until the end, there would be one
\texttt{\textless{}interjection\textgreater{}} node and one
\texttt{\textless{}continuation\textgreater{}} node following the
\texttt{\textless{}talk.text\textgreater{}} node. These would contain
details on the person who made each statement, such as their party and
electorate.

In contrast, the speech contents in XMLs from 1998 up to and including
24 March 2011 are nested differently --- there is no
\texttt{\textless{}talk.text\textgreater{}} node. Rather than this
single child-node that contains all speech content, statements are
categorized in individual child-nodes. This means that unlike our code
for parsing more current Hansards, we cannot specify a single XPath
expression such as
``\texttt{chamber.xscript//debate//speech/talk.text}'' to extract all
speeches, in their entirety, at once. This difference in nesting
structure made many components of our second script unusable for
processing transcripts preceding 10 May 2011, and required us to change
our data processing approach considerably.

Since the earlier Hansard XMLs do not have a
\texttt{\textless{}talk.text\textgreater{}} node, we found that the most
straightforward way to preserve the ordering of statements and to parse
all speech contents at once was to parse from the
\texttt{\textless{}debate\textgreater{}} element directly. The reason we
did not use its \texttt{\textless{}speech\textgreater{}} child-node is
because every speech has a unique structure of node children, and this
makes it difficult to write code for data cleaning which is
generalizable across all speeches and sitting days. The challenge with
parsing through the \texttt{\textless{}debate\textgreater{}} element is
that every piece of data stored in that element is parsed as a single
string, including all \texttt{\textless{}talk.start\textgreater{}} data,
and all nested sub-debate data. For example, the data shown in
Figure~\ref{fig-patternEX} would be parsed as a single string preceding
the speech content, like so:

\begin{quote}
\texttt{09:31:0010261Costello,\ Peter,\ MPMr\ COSTELLOCT4HigginsLPTreasurer10}
\end{quote}

\begin{figure}

{\centering \includegraphics[width=3.28125in,height=\textheight]{images/patternEX.png}

}

\caption{\label{fig-patternEX}Portion of XML file for Hansard on 12
December 2002}

\end{figure}

This was not isolated to just the beginning of speeches --- details on
individuals interjecting or commenting during speeches were also
captured this way. To separate statements correctly, we collected all of
these patterns using the \texttt{\textless{}talk.start\textgreater{}}
node, and used them to split statements wherever one of these patterns
was found. After separating the statements, we were able to remove these
patterns from the body of text. We also used this method of extracting
and later removing unwanted patterns for other pieces of data which did
not belong to the debate proceedings, such as sub-debate titles.

Once we finalized this new method of processing the data, we proceeded
with data cleaning using the same general approach as in the first two
scripts to produce the same structure of CSV output. We then worked
backwards in time and modified the code as needed for generalizability.
Throughout this process we found many transcription errors present in
the XMLs from earlier years. We fixed these manually, deferring to the
official release to ensure the correct information was filled in. Since
there were a number of transcription errors specific to the 2000s, we
chose to create a fourth script for parsing 1998 and 1999. This allowed
us to remove all the code which was needed to resolve specific
transcription errors of the 2000s, to avoid an overly long script and in
turn improving computational efficiency. As such, our fourth script is
essentially the same as the third, with the only difference being that
it has code specific to fixing transcription errors from 1998 and 1999.

\hypertarget{sec-qa}{%
\subsection{Question Time}\label{sec-qa}}

A key characteristic of the Australian parliamentary system is the
ability for the executive government to be held accountable for their
decisions. One core mechanism by which this is achieved is called
Question Time. This is a period of each sitting day in the Chamber where
members of the House can ask ministers two types of questions: questions
in writing which are written in advance, or questions without notice
which are asked verbally in the Chamber and are responded to in real
time (House of Representatives 2021). Questions without notice are
included directly in the \texttt{"chamber.xscript"} child node, with
sub-child nodes called \texttt{"question"} and \texttt{"answer"} to
differentiate the two. Questions in writing, however, are embedded in
their own child node called \texttt{"answers.to.questions"} at the end
of the XML file.

Our approach to parse the \texttt{"chamber.xscript"} speeches used in
all four scripts meant that all questions without notice content was
already parsed in order. For the first two scripts, questions and
answers were already separated onto their own rows. For the third and
fourth scripts, just as we did with the rest of the speech content, we
used those patterns of data preceding the text to separate questions and
answers. Finally, since questions in writing exist in their own child
node we were able to use the same parsing method for all scripts, which
was to extract all question and answer elements from the
\texttt{"answers.to.questions"} child node.

We then added binary flags to differentiate between questions and
answers. To do this in the first and second scripts, we separately
re-parsed question and answer content using the XPath expressions
\texttt{"chamber.xscript//question"} and
\texttt{"chamber.xscript//answer"}, added the correct question and
answer flags accordingly, and then added those flags back to the main
dataframe based on exact text matches. For the third and fourth scripts,
we made use of the fact that the patterns preceding text transcribed
under a question node were stored separately from those transcribed
under an answer node. As a result, we could readily use those patterns
to flag questions and answers correctly based on which list of patterns
it belonged to. Sometimes, questions were incorrectly transcribed under
an answer node and vice-versa, in which cases we manually corrected the
question and answer flags. For instance, we check for any statements
flagged as questions which include the phrase ``has provided the
following answer to the honourable member's question'', in which case we
re-code that statement as an answer.

The next step was to merge Question Time contents with all the debate
speech. Our method of parsing meant that everything was already in
order, so we did not have to perform any additional merging. For
questions in writing, merging this content was also straightforward due
to the fact that it is always at the end of Hansard. This means that we
could bind question in writing rows to the bottom of the main dataframe.
This approach was used for all four scripts.

\hypertarget{sec-interject}{%
\subsection{Interjections}\label{sec-interject}}

As mentioned, the text was structured and parsed in such a way that
various interjections and comments which happened during a speech were
not separated onto individual rows. This was the case across the entire
time frame of documents. We will first discuss the methodology employed
to split interjections in the first and second scripts, as it informed
our approach for the third and fourth scripts.

Below is an example of part of a speech we would need to split,
extracted from Hansard on 30 November 2021, where Bert van Manen is
interrupted by the Speaker who states that the time for members'
statements has concluded.

\begin{quote}
``Mr VAN MANEN (Forde---Chief Government Whip) (13:59): It's a great
pleasure to share with the House that Windaroo Valley State High School
has qualified for the finals of the Australian Space Design Competition,
to begin in January next year. The competition is regarded as the
premier STEM competition for high school students and is recognised by
universities around the country. The students are required to respond to
industry-level engineering and requests for tender for design and---The
SPEAKER: Order! In accordance with standing order 43, the time for
members' statements has concluded.''
\end{quote}

We want each statement on its own row with the correct name, name ID,
electorate and party information on the individual speaking. We
approached this task in a number of steps.

Once all parsed text from the XML was merged into one dataframe called
\texttt{main}, our first step was to add a \texttt{"speech\_no"}
variable. This was done to keep track of which speech each interjection,
comment, or continuation belonged to as we separated these components
onto their own rows.

The next step was to extract all the names and titles preceding these
interjections, comments and continuations. This would enable us to then
separate the speeches in the correct places using these names and titles
in combination with regular expressions, which are patterns of
characters that can be used to search bodies of text. We completed this
extraction process with a few intermediate steps, due to the large
number of name styles and interjection types that had to be accounted
for, each requiring their own unique regular expression format.

As mentioned in Section~\ref{sec-diff}, more recent years of Hansard
XMLs contain a series of child-nodes which exist to capture the
structure of interruptions in that speech. Figure~\ref{fig-xml3}
provides an example of this, where the speech was interrupted by a
comment from the Deputy Speaker, and then the member continued their
speech. Looking at the element names highlighted in blue, these
child-nodes do not contain the actual text for the interjection or
continuation --- this text is embedded within the speech above it.
However, as shown by the content highlighted in pink in
Figure~\ref{fig-xml3}, we were able to extract useful details on the
individual interjecting which we could use later. Making use of this
structure, we extracted names and information of all individuals that
were categorized within the XML as interjections. We stored this as a
dataframe called \texttt{"interject"}. We decided not to include this
data in our final database, as it is embedded in our resulting datasets
which have a flag for interjections.

\begin{figure}

{\centering \includegraphics[width=3.59375in,height=\textheight]{images/xml3.png}

}

\caption{\label{fig-xml3}Snapshot of XML structure with interjection and
continuation from 03 February 2021 Hansard}

\end{figure}

We then created lists using both the \texttt{interject} and
\texttt{main} dataframes to capture all the names of individuals who
spoke that day. We added the names of all members in a number of unique
formats, due to the frequent variation in how names are transcribed in
Hansard. When a member interjects or continues a speech, the usual form
of their name is a title followed by their first name or first initial
and/or last name. There is also variation in the capitalization of these
names.\footnote{Sometimes when someone's first name is included, only
  their last name is capitalized, while sometimes their full name is
  capitalized, or other times neither are capitalized.} Another source
of variation is in individuals with more than one first name, as
sometimes only their initial first name is written, while other times
their entire first name is written. Additionally, some surnames have
punctuation, and some surnames have specific capitalization such as
``McCormack'', where even in full capitalization, the first ``c''
remains lower case. This variation demands careful consideration when
writing regular expression patterns. In these lists we also accounted
for any general interjection statements that were not attributed to an
individual, such as ``An opposition member interjecting-''.

Having these lists enabled us to extract the names of members and their
associated titles as they exist in the text, by searching for exact
matches with regular expression patterns. We then used these extracted
names to split all the speeches, using regular expressions with
lookarounds. A lookaround can be added to a regular expression pattern
to enhance the specificity of matches. These were used to ensure that
the text was not being split in the wrong places, such as places where
members were being named in the statement of another member.

Once all interjections, comments and continuations were successfully
split onto their own rows using the lists we created, we did one final
check for any additional names that were not captured in these lists. To
do so, we searched for any remaining name matches in speech bodies with
general regular expressions and lookarounds, and separated text using
those matches when found.

We then added an order variable to the dataset based on row number, to
keep track of the order in which everything was said. The next step was
to fill the name, name ID, electorate and party variables with the
correct data for each row. We also wanted to add the gender and unique
identifier for each individual as found in the
\texttt{AustralianPoliticians} package. To do so, we created a lookup
table, which contained the unique incomplete form in which the name was
transcribed, and the corresponding full name, name ID, electorate,
party, gender, and unique ID for that individual.
Figure~\ref{fig-lookup} provides an example of this. We used the main
dataset from the \texttt{AustralianPoliticians} package in the creation
of each lookup table (Alexander and Hodgetts 2021).

\begin{figure}

{\centering \includegraphics[width=4.66667in,height=\textheight]{images/lookup_ex.png}

}

\caption{\label{fig-lookup}First 10 rows of lookup table from 19 October
2017 Hansard processing}

\end{figure}

Next, we merged our main dataframe with the lookup table to replace any
incomplete names with their full names, and to fill in any gaps with
available name ID, electorate, party, gender, and unique ID information.
Finally, we were able to add a flag for interjections. Grouping our data
by the speech number, we defined an interjection as a statement made by
anyone who is not the Speaker, the Deputy Speaker, or the member whose
turn it was to speak. Figure~\ref{fig-interject} provides an example of
a Federation Chamber proceeding with interjections. Statements made by
the member whose turn it was to speak, or by the Deputy Speaker Maria
Vamvakinou, are not flagged as interjections.

\begin{figure}

{\centering \includegraphics{images/interject_ex.png}

}

\caption{\label{fig-interject}Example of speech with interjections from
21 November 2016 Hansard}

\end{figure}

Having developed a successful methodology for splitting interjections,
we used this to inform our general approach in the third and fourth
scripts. However, the difference in data cleaning used in these scripts
necessitated some departure from the original methodology. As discussed
in Section~\ref{sec-diff}, we used string patterns extracted from
\texttt{\textless{}talk.start\textgreater{}} nodes to separate speeches.
As evident in Figure~\ref{fig-xml3},
\texttt{\textless{}talk.start\textgreater{}} nodes are nested within
\texttt{\textless{}interjection\textgreater{}} nodes, meaning that the
patterns of data from interjection statements were separated out in the
process. In other words, our approach to data cleaning in the third and
fourth scripts separated out interjections in the process. This meant
that we did not need to create lists of names and titles for which to
search in the text as we did before. However, we used the same list of
general interjection statements on which to separate as was used in the
first two scripts. We then did an additional check for statements that
may have not been separated due to how they were embedded in the XML,
and separated those out where needed.\footnote{While most statements
  were categorized in their own child-node and hence captured through
  pattern-based separation, some were not individually categorized, and
  had to be split manually in this step.}

We then proceeded to clean up speeches and fill in correct details on
the member of parliament speaking. While we used the same lookup table
approach as before, we did so in combination with another means of
filling in these details. The patterns parsed from
\texttt{\textless{}talk.start\textgreater{}} nodes contain important
data on the member of parliament making each statement. As such, we
could extract those data associated with each pattern by parsing one
element inward, using the XPath expression \texttt{"talk.start/talker"}.
We created a pattern lookup table with these data, and merged it with
the main Hansard dataframe by the first pattern detected in each
statement. Figure~\ref{fig-patterns} provides an example of that lookup
table. This approach enabled us to fill in missing data on each member
of parliament speaking using data extracted directly from the XML.
Finally, we then used the \texttt{AustralianPoliticians} dataset to fill
in other missing data, and flagged for interjections in the same manner
as before.

\begin{figure}

{\centering \includegraphics{images/pattern_lookup.png}

}

\caption{\label{fig-patterns}10 rows of pattern lookup table from 12
December 2012 Hansard processing}

\end{figure}

\hypertarget{sec-stage}{%
\subsection{Stage Directions}\label{sec-stage}}

When building our first scripts, one of the final components needed was
to separate general stage directions out from statements made by
members. Stage directions are general statements included in the
transcript to document happenings in parliament. Examples of stage
directions are ``Bill read a second time'', ``Question agreed to'', or
``Debate adjourned''. It was unclear to us from the XML and PDF who
exactly these statements were attributed to. For further clarification,
we watched portions of the video recording for some sitting days, and
noticed that while these statements are documented in Hansard, they are
not explicitly stated in parliament. For example, when the Deputy
Speaker says ``The question is that the bill be now read a second
time'', members of the House take a vote, and if the majority is in
favour, they proceed reading the bill the second time. This vote and
second reading is not explicitly transcribed, rather what is written is:
``Question agreed to. Bill read a second time''. For this reason, we
filled the name variable for these statements with ``stage direction''.
Stage directions were not flagged as interjections. These stage
directions are not defined differently from the regular debate speech in
the XML, meaning we had to manually create a list of stage directions to
separate out of the speeches. We built this list of stage directions as
we worked backwards in parsing Hansard, and took the same approach
across all four scripts.

\hypertarget{filling-missing-details}{%
\subsection{Filling Missing Details}\label{filling-missing-details}}

While we did our best to maximize the completeness of the files in our
database as they were processed in the initial four scripts, there were
still a number of rows in which details on the person speaking were
missing, or the name transcribed for that individual was in a short form
(i.e.~``Mr Abbott'' instead of ``Abbott, Tony, MP''). This was a
particularly frequent occurrence for sitting days where a member spoke
whose surname was shared by any other past or present member of
parliament, and automated filling of their details using data from the
\texttt{AustralianPoliticians} package was avoided to prevent any
incorrect detail attribution. In an effort to improve as many of these
as possible, we developed a script which identifies short-form names
belonging to people with common surnames in each CSV, looks for the full
version of that individuals name if available in that same CSV file, and
replaces the short-form name with the full name, and fills the rest of
the MP details in accordingly with data from the
\texttt{AustralianPoliticians} package. This script does the same for
anyone who does have a unique surname but is still missing the full name
form or any gender, unique ID, name ID, party or electorate details.
Each file in our database passed through this script after being
created, to ensure it is as complete as possible.

\hypertarget{data-records}{%
\section{Data Records}\label{data-records}}

Our database is available in both CSV and parquet formats. It covers all
sitting days of the House of Representatives from 02 March 1998 to 08
September 2022, so there are 1,517 files for each format. All data
records are available on the general-purpose repository Zenodo, at
https://doi.org/10.5281/zenodo.7336075 (Katz and Alexander 2023). For
each file, each row contains an individual statement, with details on
the individual speaking. For general statements transcribed as made by
``Honourable members'' for example, these variables cannot be specified.
Table~\ref{tbl-vars} provides an overview of each variable found in the
database.

\hypertarget{tbl-vars}{}
\begin{table}[H]
\caption{\label{tbl-vars}Summary and description of variables in our database }\tabularnewline

\centering
\begin{tabular}{ll}
\toprule
Variable & Description\\
\midrule
name & Name of speaker\\
order & Row number\\
speech\_no & Speech number\\
page.no & Page number statement can be found on in official Hansard\\
time.stamp & Time of statement\\
\addlinespace
name.id & Unique member identification code, based on the Parliamentary Handbook\\
electorate & Speaking member's electorate\\
party & Speaking member's party\\
in.gov & Flag for in government (1 if in government, 0 otherwise)\\
first.speech & Flag for first speech (1 if first speech, 0 otherwise)\\
\addlinespace
body & Statement text\\
fedchamb\_flag & Flag for Federation Chamber (1 if Federation Chamber, 0 if Chamber)\\
sub1\_flag & Flag for sub-debate 1 contents (1 if sub-debate 1, 0 otherwise)\\
sub2\_flag & Flag for sub-debate 2 contents (1 if sub-debate 2, 0 otherwise)\\
question & Flag for question (1 if question, 0 otherwise)\\
\addlinespace
answer & Flag for answer (1 if answer, 0 otherwise)\\
q\_in\_writing & Flag for question in writing (1 if question in writing, 0 otherwise)\\
div\_flag & Flag for division (1 if division, 0 otherwise)\\
gender & Gender of speaker\\
uniqueID & Unique identifier of speaker\\
\addlinespace
interject & Flag for interjection (1 if statement is an interjection, 0 otherwise)\\
\bottomrule
\end{tabular}
\end{table}

The \texttt{name}, \texttt{page.no}, \texttt{time.stamp},
\texttt{name.id}, \texttt{electorate}, \texttt{party}, \texttt{in.gov},
\texttt{first,speech}, and \texttt{body} variables all came directly
from the XML contents. In addition to these variables, we added a number
of flags to enable easy filtering of statements. For example, adding the
\texttt{fedchamb\_flag} provides a clear distinction between the
proceedings of the Chamber with those of the Federation Chamber. As
well, the \texttt{sub1\_flag} and \texttt{sub2\_flag} variables allow us
to keep track of where various statements are being parsed from in the
XML document. The \texttt{question}, \texttt{answer}, and
\texttt{q\_in\_writing} flags were added to identify statements
belonging to Question Time, and the nature of these statements. We also
flagged for interjections (\texttt{interject}), and the
\texttt{div\_flag} variable was added to flag when a division is called
for. The \texttt{gender} and \texttt{uniqueID} variables were added
based on the main dataset from the \texttt{AustralianPoliticians}
package. Details on the usage of \texttt{uniqueID} will be provided in
Section~\ref{sec-usage}. Further, the \texttt{speech\_no} variable
allows us to keep track of the speech number that each statement and
interjection belongs to. Having the speech number variable offers an
easy way to group statements by speech or isolate specific speeches of
interest. Finally, the \texttt{order} variable was added to maintain the
order of proceedings.

\hypertarget{technical-validation}{%
\section{Technical Validation}\label{technical-validation}}

We developed a script to perform automated tests on each file in our
database, to enhance its quality and consistency. Our first test
validates that the date specified in each file name matches the date
specified in its corresponding XML session header.\footnote{As seen in
  Figure~\ref{fig-xml1}, the first child node of the
  \texttt{\textless{}session.header\textgreater{}} element is the date.}
Every file passed this test, and we detected one discrepancy in an XML
file from 03 June 2009, where its session header contained the wrong
date. We validated that our file name and date was correct by checking
the official PDF release from that sitting day.

The second test is designed to test for duplication errors in the data,
by checking whether two lines that immediately follow each other have
the same body (i.e.~spoken content). This test detected 129 dates on
which duplicate statements were made, and one immediately follows the
other. Note that this test does not account for who is making each
statement, meaning one member repeating the words of another member
would be picked up in this test as well. We checked a sample of 40\% of
these duplicates, and manually validated that they are all repeated
statements that do exist and are transcribed closely together in that
day's XML file, and by our method should be parsed such that one of
these statements is immediately followed by the other.

When a member runs out of allotted time for their speech, Hansard
editors transcribe ``(Time expired)'' after their final word. As a means
of checking that we have separated speeches out correctly, our third
test checks that when the phrase ``(Time expired)'' exists in a body of
text, it exists at the very end. When this is not the case, we know that
we have missed the separation of the next statement onto its own row,
and could fix this accordingly.

The remaining tests focus on the members present on each sitting day.
Our fourth test checks that there is one unique party and electorate
attributed to each individual on each sitting day. As we parsed Hansard
further back in time, we found a number of cases where an individual was
associated with the wrong electorate or party due to transcription
errors. When we found these data errors we corrected them based on the
official release. This test provides us with an automated way to catch
these errors and correct them at scale.

Next, we test that the unique name identification code attributed to
each individual is found in the Australian Parliamentary Handbook. We do
so using the \texttt{ausPH} package. This test serves as another means
to correct for transcription errors, this time in the case of name IDs.
We found and corrected for a number of common name ID transcription
errors detected by this test, such as a capital letter ``O'' in place of
a zero.

Our sixth test checks that on any given sitting day, the individuals
identified are alive. To do so, we utilized the main dataset from the
\texttt{AustralianPoliticians} package which contains the birth and
where applicable death dates for every politician. This test confirmed
that all members who are detected to speak on each sitting day are not
deceased.

Finally, our seventh test validates that all individuals speaking are
active members of parliament on that particular day. We use the
\texttt{mps} dataset from the \texttt{AustralianPoliticians} package
which has the dates of each Member's duration in parliament. Using these
dates, we check that each person speaking on each sitting day is in fact
a member of parliament on that day.

\hypertarget{next-steps}{%
\section{Next Steps}\label{next-steps}}

While our database makes a substantial contribution to the field, there
are a few areas for future development.

One important contribution would be to take advantage of similar records
that exist for the Senate over this same period and add them to our
database. The main challenge of this would be accounting for different
parliamentarians, but the broad structures would be similar, making many
elements of our code and methodology applicable to this work.

Earlier House of Representatives records also exist, and would be
valuable to include in future versions of our database. In some cases
however it is difficult to get XML records for certain dates, and the
earlier decades would require more cleaning and likely additional script
development to account for changes in structural formatting.

Having completed this work, we recognize components of our scripts which
could be developed in a more computationally efficient manner such as
our approach to splitting interjections and filling in MP details. In
future, it would be beneficial to review our methodological approach to
these steps, and consider approaches which are more concise where
appropriate.

Finally, Hansard records contain a number of statements throughout that
were not explicitly spoken in parliament. For example, the text in
Figure~\ref{fig-content_ex} beneath the bold subtitle ``Report from
Federation Chamber'' is not attributed to a particular member and was
not spoken aloud during the parliamentary proceedings. In contrast, the
text which comes afterward was spoken aloud in parliament by the Speaker
at 14:01. Generally, unspoken text is nested under
\texttt{\textless{}debate.text\textgreater{}} or
\texttt{\textless{}subdebate.text\textgreater{}} nodes, while spoken
text is categorized under \texttt{\textless{}speech\textgreater{}}.
Importantly, in our work we focus primarily on parsing spoken speech
content (i.e.~what is nested within
\texttt{\textless{}speech\textgreater{}} nodes), such as what is stated
by the Speaker in Figure~\ref{fig-content_ex}. For this reason, many
Hansard transcriptions which are not spoken text may not be captured in
our database. In future work, capturing this unspoken content in our
database would be a priority.

\begin{figure}

{\centering \includegraphics[width=1.88542in,height=\textheight]{images/content_ex.png}

}

\caption{\label{fig-content_ex}Snapshot from page 7129 of the PDF file
for Hansard on 26 March 2013}

\end{figure}

\hypertarget{sec-usage}{%
\section{Usage Notes}\label{sec-usage}}

To enhance the usability of our database, we added a \texttt{uniqueID}
variable to each CSV file. This serves as a unique identifier for each
speaking Member, and comes from the \texttt{uniqueID} variable present
within data from both the \texttt{AustralianPoliticians} and
\texttt{AustralianElections} R packages (Alexander and Hodgetts 2021;
Alexander 2019). By including this variable, one can readily integrate
our data records with those available in these two packages.

Further, the \texttt{name.id} variable found in each CSV is another
unique identifier for each member of parliament. This variable was
parsed directly from the Hansard XML files, and can be found in the
Australian Parliamentary Handbook. As such, our data records can be
integrated with those from the \texttt{ausPH} package which provides
datasets for contents of the Australian Parliamentary Handbook (Leslie
2021). This will allow for convenient extraction of further details on
each member of parliament in a tidy, ready to analyze format.

\hypertarget{code-availability}{%
\section{Code Availability}\label{code-availability}}

The code written to build this database is available on the GitHub
repository associated with this paper:
https://github.com/lindsaykatz/hansard-proj. All scripts were created
using R software (R Core Team 2022). The core packages used to develop
these scripts are: the \texttt{XML} package by Temple Lang (2022), the
\texttt{xml2} package by Wickham, Hester, and Ooms (2021), the
\texttt{tidyverse} R packages by Wickham et al. (2019), the
\texttt{AustralianPoliticians} package by Alexander and Hodgetts (2021),
and the \texttt{ausPH} package by Leslie (2021). \texttt{XML} and
\texttt{xml2} were used for parsing the XML documents,
\texttt{AustralianPoliticians} and \texttt{ausPH} were used for cleaning
up and filling in member details in the datasets, and \texttt{tidyverse}
packages were used in all steps, for tidy wrangling of data.

\hypertarget{acknowledgements}{%
\section{Acknowledgements}\label{acknowledgements}}

We thank Kristine Villaluna, Monica Alexander, and Jack Stephenson for
helpful comments.

\hypertarget{author-contributions}{%
\section{Author Contributions}\label{author-contributions}}

LK developed and implemented the code to obtain, create, and test, the
datasets and wrote the first draft of the paper. RA conceptualized and
designed the study, and contributed writing. Both authors approved the
final version.

\hypertarget{competing-interests}{%
\section{Competing Interests}\label{competing-interests}}

The authors declare no competing interests.

\hypertarget{references}{%
\section*{References}\label{references}}
\addcontentsline{toc}{section}{References}

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\vadjust pre{\hypertarget{ref-auselect}{}}%
Alexander, Rohan. 2019. \emph{AustralianElections: Datasets on
Australian Elections}.
\url{https://github.com/RohanAlexander/AustralianElections}.

\leavevmode\vadjust pre{\hypertarget{ref-alexander2021}{}}%
Alexander, Rohan, and Monica Alexander. 2021. {``The Increased Effect of
Elections and Changing Prime Ministers on Topics Discussed in the
Australian Federal Parliament Between 1901 and 2018.''} \emph{arXiv
Preprint arXiv:2111.09299}.

\leavevmode\vadjust pre{\hypertarget{ref-auspol}{}}%
Alexander, Rohan, and Paul A. Hodgetts. 2021.
\emph{AustralianPoliticians: Provides Datasets about Australian
Politicians}.
\url{https://CRAN.R-project.org/package=AustralianPoliticians}.

\leavevmode\vadjust pre{\hypertarget{ref-lipad}{}}%
Beelen, Kaspar, Timothy Alberdingk Thijm, Christopher Cochrane, Kees
Halvemaan, Graeme Hirst, Michael Kimmins, Sander Lijbrink, et al. 2017.
{``Digitization of the Canadian Parliamentary Debates.''} \emph{Canadian
Journal of Political Science/Revue Canadienne de Science Politique} 50
(3): 849--64.

\leavevmode\vadjust pre{\hypertarget{ref-fraussen2018}{}}%
Fraussen, Bert, Timothy Graham, and Darren R Halpin. 2018. {``Assessing
the Prominence of Interest Groups in Parliament: A Supervised Machine
Learning Approach.''} \emph{The Journal of Legislative Studies} 24 (4):
450--74.

\leavevmode\vadjust pre{\hypertarget{ref-australia2018house}{}}%
House of Representatives. 2018. Edited by D. R. Elder and P. E. Fowler.
\emph{{House of Representatives Practice}}. 7th ed. {Australian
Government - Department of the House of Representatives}.

\leavevmode\vadjust pre{\hypertarget{ref-house2021}{}}%
---------. 2021. \emph{{A window on the house: Practices and procedures
relating to question time}}. {Parliament of Australia}.

\leavevmode\vadjust pre{\hypertarget{ref-katz_2023}{}}%
Katz, Lindsay, and Rohan Alexander. 2023. {``{A new, comprehensive
database of all proceedings of the Australian Parliamentary Debates
(1998-2022)}.''} Zenodo. \url{https://doi.org/10.5281/zenodo.7785767}.

\leavevmode\vadjust pre{\hypertarget{ref-Leslie_2021_HoR_Aus}{}}%
Leslie, Patrick. 2021. \emph{Members of the Australian House of
Representatives (1945-2019)}.
\url{https://github.com/palesl/AustralianHouseOfRepresentatives}.

\leavevmode\vadjust pre{\hypertarget{ref-r_software}{}}%
R Core Team. 2022. \emph{R: A Language and Environment for Statistical
Computing}. Vienna, Austria: R Foundation for Statistical Computing.
\url{https://www.R-project.org/}.

\leavevmode\vadjust pre{\hypertarget{ref-rasiah2010framework}{}}%
Rasiah, Parameswary. 2010. {``A Framework for the Systematic Analysis of
Evasion in Parliamentary Discourse.''} \emph{Journal of Pragmatics} 42
(3): 664--80.

\leavevmode\vadjust pre{\hypertarget{ref-salisbury2011}{}}%
Salisbury, Christopher. 2011. {``'Mr Speaker, i Withdraw...': Standards
of (Mis) Behaviour in the Queensland, Western Australian and
Commonwealth Parliaments Compared via Online Hansard.''}
\emph{Australasian Parliamentary Review} 26 (1): 166--77.

\leavevmode\vadjust pre{\hypertarget{ref-sherratt}{}}%
Sherratt, Tim. 2016. {``Documentation: Historic Hansard.''}
\emph{Historic Hansard}.
\url{http://timsherratt.org/digital-heritage-handbook/docs/historic-hansard/\#}.

\leavevmode\vadjust pre{\hypertarget{ref-XML_pkg}{}}%
Temple Lang, Duncan. 2022. \emph{XML: Tools for Parsing and Generating
XML Within r and s-Plus}. \url{https://CRAN.R-project.org/package=XML}.

\leavevmode\vadjust pre{\hypertarget{ref-vice2017history}{}}%
Vice, John, and Stephen Farrell. 2017. \emph{The History of Hansard}.
House of Lords Library; House of Lords Hansard.

\leavevmode\vadjust pre{\hypertarget{ref-Wickham2014}{}}%
Wickham, Hadley. 2014. {``Tidy Data.''} \emph{Journal of Statistical
Software} 59 (10). \url{https://doi.org/10.18637/jss.v059.i10}.

\leavevmode\vadjust pre{\hypertarget{ref-tidyverse}{}}%
Wickham, Hadley, Mara Averick, Jennifer Bryan, Winston Chang, Lucy
D'Agostino McGowan, Romain François, Garrett Grolemund, et al. 2019.
{``Welcome to the {tidyverse}.''} \emph{Journal of Open Source Software}
4 (43): 1686. \url{https://doi.org/10.21105/joss.01686}.

\leavevmode\vadjust pre{\hypertarget{ref-xml2_pkg}{}}%
Wickham, Hadley, Jim Hester, and Jeroen Ooms. 2021. \emph{Xml2: Parse
XML}. \url{https://CRAN.R-project.org/package=xml2}.

\end{CSLReferences}

\end{document}