### 2023-07-03 scripts/13-summary_stats.Rmd
rename scripts and update readme

---
title: "summary_stats"
author: "Lindsay Katz"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
```

```{r}
library(tidyverse)
library(arrow)

# read in corpus
corpus <- read_parquet("/Volumes/Verbatim/v4/hansard_1998_to_2022-parquet/hansard_corpus_1998_to_2022.parquet")

# read in party table
party_table_final <- read_csv("../additional_data/PartyFacts_map.csv", show_col_types = F)
```

## Total distinct speeches by party facts ID

```{r}
# only for those parties with a party facts ID
s_by_party <- corpus |> select(date, speech_no, partyfacts_id, party) |> 
  mutate(party = as.character(ifelse(str_detect(party, "^Ind"), "IND", as.character(party)))) |> 
  group_by(date, partyfacts_id) |> 
  summarise(n_distinct_speeches = n_distinct(speech_no, na.rm=T)) |> 
  group_by(partyfacts_id) |> 
  summarise(total_distinct_speeches = sum(n_distinct_speeches)) |> 
  filter(!is.na(partyfacts_id)) |> 
  left_join(party_table_final |> 
              select(-party_abb_hansard, -party_abb_auspol) |> 
              distinct() |> 
              # b/c Centre Alliance is current name and they have the same party facts ID
              filter(party_name_auspol!="Nick Xenophon Team"), 
            by="partyfacts_id")

# rename vars
s_by_party |> select(party_name_auspol, everything()) |> 
  arrange(desc(total_distinct_speeches)) |> 
  rename("Party Name" = party_name_auspol,
         "PartyFacts ID" = partyfacts_id,
         "# Distinct Speeches" = total_distinct_speeches) |> 
  kableExtra::kable()
```

## Total distinct speeches by Australian Politicians party name

```{r}
corpus |> select(date, speech_no, party) |> 
  filter(!is.na(party)) |> 
  left_join(party_table_final, by=c("party" = "party_abb_hansard")) |> 
  group_by(date, party_abb_auspol, party_name_auspol) |> 
  summarise(n_distinct_speeches = n_distinct(speech_no, na.rm=T)) |> 
  group_by(party_abb_auspol, party_name_auspol) |> 
  summarise(total_distinct_speeches = sum(n_distinct_speeches))

# only for those parties with a party facts ID
s_by_party <- corpus |> select(date, speech_no, partyfacts_id, party) |> 
  mutate(party = as.character(ifelse(str_detect(party, "^Ind"), "IND", as.character(party)))) |> 
  group_by(date, partyfacts_id) |> 
  summarise(n_distinct_speeches = n_distinct(speech_no, na.rm=T)) |> 
  group_by(partyfacts_id) |> 
  summarise(total_distinct_speeches = sum(n_distinct_speeches)) |> 
  filter(!is.na(partyfacts_id)) |> 
  left_join(party_table_final |> 
              select(-party_abb_hansard, -party_abb_auspol) |> 
              distinct() |> 
              # b/c Centre Alliance is current name and they have the same party facts ID
              filter(party_name_auspol!="Nick Xenophon Team"), 
            by="partyfacts_id")

# rename vars
s_by_party |> select(party_name_auspol, everything()) |> 
  arrange(desc(total_distinct_speeches)) |> 
  rename("Party Name" = party_name_auspol,
         "PartyFacts ID" = partyfacts_id,
         "# Distinct Speeches" = total_distinct_speeches) |> 
  kableExtra::kable()
```

## Total daily distinct speeches

```{r}
s_by_day <- corpus %>% select(date, speech_no, fedchamb_flag) %>%
  group_by(date, fedchamb_flag) %>% 
  summarise(n_distinct_speeches = n_distinct(speech_no, na.rm=T))

# plot
ggplot(s_by_day, aes(x=as.Date(date), y=n_distinct_speeches)) + 
  geom_point() +
  geom_smooth(method="loess") +
  facet_wrap(~fedchamb_flag, labeller = as_labeller(c(`0` = "Chamber", `1` = "Federation Chamber"))) +
  labs(x = "date", y = "# distinct speeches")
```

## Total yearly distinct speeches

```{r}
s_by_year <- corpus %>% select(date, speech_no, fedchamb_flag) %>%
  group_by(date, fedchamb_flag) %>% 
  summarise(n_distinct_speeches = n_distinct(speech_no, na.rm=T)) %>% 
  mutate(year = str_extract(date, "^\\d{4}")) %>% 
  group_by(year, fedchamb_flag) %>% 
  summarise(total_speeches = sum(n_distinct_speeches))

# plot
s_by_year |> filter(year!="2022") |> 
ggplot(aes(x=as.numeric(year), y=total_speeches)) + 
  geom_point() +
  geom_smooth(method="loess") +
  facet_wrap(~fedchamb_flag, labeller = as_labeller(c(`0` = "Chamber", `1` = "Federation Chamber"))) +
  labs(x = "year", y = "# distinct speeches")
```

## Distinct speeches by electorate

```{r}
s_by_electorate <- corpus |> group_by(date, electorate) |> 
  summarise(n_distinct_speeches = n_distinct(speech_no, na.rm=T)) |> 
  group_by(electorate) |> 
  summarise(total_speeches_electorate = sum(n_distinct_speeches)) |> 
  filter(!is.na(electorate))

# look at avg
s_by_electorate |> summarise(avg = mean(total_speeches_electorate))

# look at top 10
s_by_electorate |> arrange(desc(total_speeches_electorate)) |> slice(1:10)

# look at bottom 10
s_by_electorate |> arrange(total_speeches_electorate) |> slice(1:10)
```

## Annual proportion of interjections, by gender

```{r}
### INTERJECTIONS BY GENDER, ANNUALLY ###
interject_yr_sex <- corpus %>% select(date, gender, interject, fedchamb_flag) %>% 
  mutate(year = str_extract(date, "^\\d{4}"),
         gender = as.character(gender),
         gender = ifelse(is.na(gender), "unknown", gender)) %>% 
  group_by(year, gender, fedchamb_flag, interject) %>% 
  summarise(n = n()) |> 
  group_by(year, fedchamb_flag) |> 
  mutate(total_rows = sum(n)) |> 
  filter(interject==1) |> 
  group_by(year, gender, fedchamb_flag) |> 
  summarise(prop_interject = 100*n/total_rows) %>% 
  ungroup()

# plot
# remove 2022 b/c incomplete
interject_yr_sex |> filter(year!="2022") |> 
ggplot(aes(x=as.numeric(year), y=prop_interject, color=as.factor(gender))) +
  geom_point() +
  geom_smooth() +
  facet_wrap(~fedchamb_flag, labeller = as_labeller(c(`0` = "Chamber", `1` = "Federation Chamber"))) +
  labs(y = "Percent Interjections", x = "Year", color = "Gender") +
  theme(legend.position="bottom")
```

## Daily proportion of interjections, by gender

```{r}
### INTERJECTIONS BY GENDER, DAILY ###
interject_day_sex <- corpus %>% select(date, gender, interject, fedchamb_flag) %>% 
  mutate(gender = as.character(gender),
         gender = ifelse(is.na(gender), "unknown", gender)) %>% 
  group_by(date, gender, fedchamb_flag, interject) %>% 
  summarise(n = n()) %>% 
  group_by(date, fedchamb_flag) |> 
  mutate(total_rows = sum(n)) |> 
  filter(interject==1) |> 
  group_by(date, gender, fedchamb_flag) |> 
  summarise(prop_interject = 100*n/total_rows) %>% 
  ungroup()

# plot
ggplot(interject_day_sex, aes(x=as.Date(date), y=prop_interject, color=as.factor(gender))) +
  geom_point(alpha=0.25) +
  geom_smooth(method="loess") +
  facet_wrap(~fedchamb_flag, labeller = as_labeller(c(`0` = "Chamber", `1` = "Federation Chamber")), scales = "free_y") +
  labs(y = "Percent Interjections", x = "Date", color = "Gender") +
  theme(legend.position="bottom")
```

## Daily word count

```{r}
## DAILY WORD COUNT ##
daily_word_count <- corpus |> 
  group_by(date) |> 
  summarise(word_count = sum(str_count(body), na.rm = T))

daily_word_count |> arrange(word_count)

daily_word_count |> filter(is.na(word_count))

# plot
ggplot(daily_word_count, aes(x=as.Date(date), y=word_count)) +
  geom_point(alpha = 0.5) +
  geom_smooth() +
  labs(x="Date", y="Word Count") +
  theme_bw()

ggsave("../paper/word_count.png", width=8, height=5)
```

## Daily number of unique names

```{r}
## DAILY # UNIQUE NAMES ##
daily_unique_names <- corpus |> 
  select(date, name, fedchamb_flag) |> 
  filter(name!="business start" & name!="stage direction") |> 
  group_by(date, name, fedchamb_flag) |> 
  distinct() |> 
  group_by(date, fedchamb_flag) |> 
  summarise(n=n())

# plot
ggplot(daily_unique_names, aes(x=as.Date(date), y = n, color=fedchamb_flag, fill=fedchamb_flag)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method="loess") +
  labs(x="Date", y="Number of unique names", color=NULL, fill=NULL) +
  scale_color_discrete(labels=c("Chamber", "Federation Chamber")) +
  scale_fill_discrete(labels=c("Chamber", "Federation Chamber")) +
  theme_bw()

ggsave("n_unique_names.png", path = "../paper/", width=8, height = 5)
```

## Annual number of divisions

```{r}
corpus |> select(date, div_flag) |> 
  mutate(year = str_extract(date, "^\\d{4}")) |> 
  filter(div_flag==1) |> 
  group_by(year) |> 
  summarise(n_divs = n())

# not sure why the numbers are so low for 2012 onward
```

## Meeting per year - chamber and federation chamber

```{r}
meeting_per_year <- corpus |> select(date, fedchamb_flag) |>
  mutate(year = str_extract(date, "^\\d{4}")) |> 
  group_by(year, fedchamb_flag) |> 
  summarise(n_sittings = n_distinct(date))
```

These match well with the numbers [here](https://www.aph.gov.au/-/media/02_Parliamentary_Business/22_Chamber_Documents/224_Statistics/House_of_Representatives/Statistics_Historical/sittings_statistics.pdf?la=en&hash=5B5E7E9518C4D7A416D0846FE0D44E063A4AB3FD)

```{r}
sitting_stats <- read_csv("~/Desktop/RA/sitting_stats.txt", col_names = FALSE)

sitting_stats <- sitting_stats |> select(X1, X3, X5) |> 
  rename(year = X1,
         chamb = X3,
         fed_chamb = X5) |> 
  pivot_longer(chamb:fed_chamb, names_to = "fedchamb_flag", values_to = "official_stats")

# find years where there is a discrepancy in # of meetings
# care more about chamber b/c often fedchamb is wrongly transcribed so there are bound to be discrepancies
meeting_per_year |> mutate(fedchamb_flag = case_when(fedchamb_flag == 0 ~"chamb",
                                                     fedchamb_flag == 1 ~"fed_chamb"),
                           year = as.numeric(year)) |> 
  left_join(sitting_stats, by=c("year", "fedchamb_flag")) |> 
  filter(year!=2022) |> 
  mutate(diff = abs(n_sittings-official_stats)) |> 
  filter(diff!=0) |> 
  arrange(desc(diff))
```

## Q in writing and Q without notice per year

```{r}
corpus |> select(date, question, q_in_writing) |>
  mutate(year = str_extract(date, "^\\d{4}")) |> 
  filter(question==1) |> 
  group_by(year, q_in_writing) |> 
  summarise(n())
```

https://www.aph.gov.au/-/media/02_Parliamentary_Business/22_Chamber_Documents/224_Statistics/House_of_Representatives/Statistics_Historical/questions_statistics.pdf?la=en&hash=A3B9F79402CC7D15B576EDAF1BC28259CDEC9553