### 2023-07-03 scripts/13-summary_stats.Rmd
summary stats

---
title: "summary_stats"
author: "Lindsay Katz"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
```

```{r}
library(tidyverse)
library(arrow)

# read in corpus
corpus <- read_parquet("/Volumes/Verbatim/v4/hansard-corpus/hansard_corpus_1998_to_2022.parquet")

# read in party table
party_table_final <- read_csv("../additional_data/PartyFacts_map.csv", show_col_types = F)
```

## Total distinct speeches by party facts ID

```{r}
# only for those parties with a party facts ID
s_by_party <- corpus |> select(date, speech_no, partyfacts_id, party) |> 
  mutate(party = as.character(ifelse(str_detect(party, "^Ind"), "IND", as.character(party)))) |> 
  group_by(date, partyfacts_id) |> 
  summarise(n_distinct_speeches = n_distinct(speech_no, na.rm=T)) |> 
  group_by(partyfacts_id) |> 
  summarise(total_distinct_speeches = sum(n_distinct_speeches)) |> 
  filter(!is.na(partyfacts_id)) |> 
  left_join(party_table_final |> 
              select(-party_abb_hansard, -party_abb_auspol) |> 
              distinct() |> 
              # b/c Centre Alliance is current name and they have the same party facts ID
              filter(party_name_auspol!="Nick Xenophon Team"), 
            by="partyfacts_id")

# rename vars
s_by_party |> select(party_name_auspol, everything()) |> 
  arrange(desc(total_distinct_speeches)) |> 
  rename("Party Name" = party_name_auspol,
         "PartyFacts ID" = partyfacts_id,
         "# Distinct Speeches" = total_distinct_speeches) |> 
  kableExtra::kable()
```

## Total distinct speeches by Australian Politicians party name

```{r}
### need to confirm party abbreviation mapping!!!!!
s_by_party_name <- corpus |> select(date, speech_no, party) |> 
  filter(!is.na(party)) |> 
  left_join(party_table_final, by=c("party" = "party_abb_hansard")) |> 
  group_by(date, party_abb_auspol, party_name_auspol) |> 
  summarise(n_distinct_speeches = n_distinct(speech_no, na.rm=T)) |> 
  group_by(party_abb_auspol, party_name_auspol) |> 
  summarise(total_distinct_speeches = sum(n_distinct_speeches)) |> 
  select(party_name_auspol, everything()) |> 
  arrange(desc(total_distinct_speeches)) |> 
  filter(!is.na(party_abb_auspol)) |> 
  rename("Party Name" = party_name_auspol,
         "Party Abbreviation" = party_abb_auspol,
         "Total # of Speeches" = total_distinct_speeches)

write_csv(s_by_party_name, "../paper/s_by_party_name.csv")
```

```{r}
parties_auspol <- AustralianPoliticians::get_auspol('allbyparty')

parties_auspol |> select(partyAbbrev, partyName) |> distinct()
```

## Total daily distinct speeches

```{r}
s_by_day <- corpus %>% select(date, speech_no, fedchamb_flag) %>%
  group_by(date, fedchamb_flag) %>% 
  summarise(n_distinct_speeches = n_distinct(speech_no, na.rm=T))

# plot
s_by_day |> filter(n_distinct_speeches!=0) |> 
  ggplot(aes(x=as.Date(date), y=n_distinct_speeches, color=fedchamb_flag, fill=fedchamb_flag)) + 
  geom_point(alpha=0.5) +
  geom_smooth() +
  #facet_wrap(~fedchamb_flag, labeller = as_labeller(c(`0` = "Chamber", `1` = "Federation Chamber"))) +
  labs(x = "Date", y = "Number of speeches", color=NULL, fill=NULL) +
  theme_bw() +
  scale_color_discrete(labels=c("Chamber", "Federation Chamber")) +
  scale_fill_discrete(labels=c("Chamber", "Federation Chamber"))

ggsave("../paper/n_daily_speeches.png", width=8, height=5)
```

## Total yearly distinct speeches

```{r}
s_by_year <- corpus %>% select(date, speech_no, fedchamb_flag) %>%
  group_by(date, fedchamb_flag) %>% 
  summarise(n_distinct_speeches = n_distinct(speech_no, na.rm=T)) %>% 
  mutate(year = str_extract(date, "^\\d{4}")) %>% 
  group_by(year, fedchamb_flag) %>% 
  summarise(total_speeches = sum(n_distinct_speeches))

# plot
s_by_year |> filter(year!="2022") |> 
ggplot(aes(x=as.numeric(year), y=total_speeches)) + 
  geom_point() +
  geom_smooth(method="loess") +
  facet_wrap(~fedchamb_flag, labeller = as_labeller(c(`0` = "Chamber", `1` = "Federation Chamber"))) +
  labs(x = "year", y = "# distinct speeches")
```

## Distinct speeches by electorate

```{r}
s_by_electorate <- corpus |> group_by(date, electorate) |> 
  summarise(n_distinct_speeches = n_distinct(speech_no, na.rm=T)) |> 
  group_by(electorate) |> 
  summarise(total_speeches_electorate = sum(n_distinct_speeches)) |> 
  filter(!is.na(electorate))

# look at avg
s_by_electorate |> summarise(avg = mean(total_speeches_electorate))

# look at top 10
s_by_electorate |> arrange(desc(total_speeches_electorate)) |> slice(1:10)

# look at bottom 10
s_by_electorate |> arrange(total_speeches_electorate) |> slice(1:10)
```

## Annual proportion of interjections, by gender

```{r}
### INTERJECTIONS BY GENDER, ANNUALLY ###
interject_yr_sex <- corpus %>% select(date, gender, interject, fedchamb_flag) %>% 
  mutate(year = str_extract(date, "^\\d{4}"),
         gender = as.character(gender),
         gender = ifelse(is.na(gender), "unknown", gender)) %>% 
  group_by(year, gender, fedchamb_flag, interject) %>% 
  summarise(n = n()) |> 
  group_by(year, fedchamb_flag) |> 
  mutate(total_rows = sum(n)) |> 
  filter(interject==1) |> 
  group_by(year, gender, fedchamb_flag) |> 
  summarise(prop_interject = 100*n/total_rows) %>% 
  ungroup()

# plot
# remove 2022 b/c incomplete
interject_yr_sex |> filter(year!="2022") |> 
ggplot(aes(x=as.numeric(year), y=prop_interject, color=as.factor(gender))) +
  geom_point() +
  geom_smooth() +
  facet_wrap(~fedchamb_flag, labeller = as_labeller(c(`0` = "Chamber", `1` = "Federation Chamber"))) +
  labs(y = "Percent Interjections", x = "Year", color = "Gender") +
  theme(legend.position="bottom") +
  theme_bw()
```

## Daily proportion of interjections, by gender

```{r}
### INTERJECTIONS BY GENDER, DAILY ###
interject_day_sex <- corpus %>% select(date, gender, interject, fedchamb_flag) %>% 
  mutate(gender = as.character(gender),
         gender = ifelse(is.na(gender), "unknown", gender)) %>% 
  group_by(date, gender, fedchamb_flag, interject) %>% 
  summarise(n = n()) %>% 
  group_by(date, fedchamb_flag) |> 
  mutate(total_rows = sum(n)) |> 
  filter(interject==1) |> 
  group_by(date, gender, fedchamb_flag) |> 
  summarise(prop_interject = 100*n/total_rows) %>% 
  ungroup()

# plot
ggplot(interject_day_sex, aes(x=as.Date(date), y=prop_interject, color=as.factor(gender))) +
  geom_point(alpha=0.25) +
  geom_smooth(method="loess") +
  facet_wrap(~fedchamb_flag, labeller = as_labeller(c(`0` = "Chamber", `1` = "Federation Chamber")), scales = "free_y") +
  labs(y = "Percent Interjections", x = "Date", color = "Gender") +
  theme(legend.position="bottom") +
  theme_bw()
```

## Daily word count

```{r}
## DAILY WORD COUNT ##
daily_word_count <- corpus |> 
  group_by(date) |> 
  summarise(word_count = sum(str_count(body), na.rm = T))

daily_word_count |> arrange(word_count)

daily_word_count |> filter(is.na(word_count))

# plot
ggplot(daily_word_count, aes(x=as.Date(date), y=word_count)) +
  geom_point(alpha = 0.5) +
  geom_smooth() +
  labs(x="Date", y="Word Count") +
  theme_bw()

ggsave("../paper/word_count.png", width=8, height=5)
```

## Daily number of unique names

```{r}
## DAILY # UNIQUE NAMES ##
daily_unique_names <- corpus |> 
  select(date, name, fedchamb_flag) |> 
  filter(name!="business start" & name!="stage direction") |> 
  group_by(date, name, fedchamb_flag) |> 
  distinct() |> 
  group_by(date, fedchamb_flag) |> 
  summarise(n=n())

# plot
ggplot(daily_unique_names, aes(x=as.Date(date), y = n, color=fedchamb_flag, fill=fedchamb_flag)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method="loess") +
  labs(x="Date", y="Number of unique names", color=NULL, fill=NULL) +
  scale_color_discrete(labels=c("Chamber", "Federation Chamber")) +
  scale_fill_discrete(labels=c("Chamber", "Federation Chamber")) +
  theme_bw()

ggsave("n_unique_names.png", path = "../paper/", width=8, height = 5)
```

## Annual number of divisions

```{r}
corpus |> select(date, div_flag) |> 
  mutate(year = str_extract(date, "^\\d{4}")) |> 
  filter(div_flag==1) |> 
  group_by(year) |> 
  summarise(n_divs = n())

# not sure why the numbers are so low for 2012 onward
```

## Meeting per year - chamber and federation chamber

```{r}
meeting_per_year <- corpus |> select(date, fedchamb_flag) |>
  mutate(year = str_extract(date, "^\\d{4}")) |> 
  group_by(year, fedchamb_flag) |> 
  summarise(n_sittings = n_distinct(date))
```

Let's compare to the numbers [here](https://www.aph.gov.au/-/media/02_Parliamentary_Business/22_Chamber_Documents/224_Statistics/House_of_Representatives/Statistics_Historical/sittings_statistics.pdf?la=en&hash=5B5E7E9518C4D7A416D0846FE0D44E063A4AB3FD).

```{r}
sitting_stats <- read_csv("~/Desktop/RA/sitting_stats.txt", col_names = FALSE)

# rename select vars based on columns in pdf linked above
sitting_stats <- sitting_stats |> select(X1, X3, X5) |> 
  rename(year = X1,
         chamb = X3,
         fed_chamb = X5) |> 
  pivot_longer(chamb:fed_chamb, names_to = "fedchamb_flag", values_to = "official_stats")

# find years where there is a discrepancy in # of meetings
# care more about chamber b/c often fedchamb is wrongly transcribed so there are bound to be discrepancies
# remove 2022 b/c we don't have complete year's data
meeting_per_year |> mutate(fedchamb_flag = case_when(fedchamb_flag == 0 ~"chamb",
                                                     fedchamb_flag == 1 ~"fed_chamb"),
                           year = as.numeric(year)) |> 
  left_join(sitting_stats, by=c("year", "fedchamb_flag")) |> 
  filter(year!=2022) |> 
  mutate(diff = abs(n_sittings-official_stats)) |> 
  filter(diff!=0) |> 
  arrange(desc(diff))
```
I was able to manually check the discrepancies for the Chamber sittings, and found that there are two days in 1998 (May 6th and 7th) and two days in 2004 (February 10th and May 12th) where the House sat in the Chamber, but there is no XML transcript for Hansard available. The Federation Chamber discrepancies may be due to transcription errors - that is, Federation Chamber proceedings being nested in the Chamber proceedings node in the XML.

```{r, eval=F, echo=F}
## Q in writing and Q without notice per year
corpus |> select(date, question, q_in_writing) |>
  mutate(year = str_extract(date, "^\\d{4}")) |> 
  filter(question==1) |> 
  group_by(year, q_in_writing) |> 
  summarise(n())

#https://www.aph.gov.au/-/media/02_Parliamentary_Business/22_Chamber_Documents/224_Statistics/House_of_Representatives/Statistics_Historical/questions_statistics.pdf?la=en&hash=A3B9F79402CC7D15B576EDAF1BC28259CDEC9553
```